---
title: "Processing LMA and LDMC from TRY for PVCurves"
author: "LDL Anderegg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



```{r import TRY data}
require(rtry)

trydatall <- rtry_import("data/TRY_Traits/27764.txt")
lookall <- rtry_explore(trydatall,DataID, DataName, TraitID, TraitName, sortBy = TraitID)
View(lookall)

```
Ancillary data to keep:
193 Vegetation type / Biome
59 Latitude
60 Longitude
61 Altitude
62 MAT
80 MAP
113 Reference/source
6601 Sampling date
213 Number of replicates
327 Treatment: Exposition
413 Plant developmental status / plant age / maturity / plant life stage
846 Family
```{r test}
# testing out James' code
Trait = data.frame(trydatall$ObsDataID,trydatall$AccSpeciesName, trydatall$StdValue, trydatall$DataName, trydatall$TraitID, trydatall$ErrorRisk)
colnames(Trait) <- c('DataID','FullName','Value','DataName','TraitID','ErrorRisk')
Trait = Trait[which(Trait$TraitID == 3117),]
datatypes = unique(Trait$DataName)
keeps = c("Leaf mass per area (LMA)")
keeps = c("SLA: undefined if petiole in- or excluded")
Trait = Trait[which(Trait$DataName %in% keeps),]
Trait = Trait[which(!is.na(Trait$Value)),]
Trait = Trait[which(Trait$FullName != "unknown"),]
Trait = Trait[which(Trait$FullName != "Coffea arabica" & Trait$FullName != "Glycine max"),]
keep = grepl('^\\w+\\s\\w+$', Trait$FullName)
Trait = Trait[keep,]
Trait <- Trait[which(Trait$ErrorRisk<5),]
Trait <- Trait %>% distinct(FullName, Value)

```

``` {r clean TRY, include=F}
trydata <- rtry_select_row(trydatall, TraitID>0 | DataID %in% c(59,60,61,62,80,113,193,6601,213,327,413,846))
#TraitID <- unique(lookall$TraitID[which(lookall$TraitID>0)])[-11] # get my trait #s, less NA
#SLA.to.exclude <- lookall$DataID[which(lookall$TraitID>0 & lookall$TraitID %in% c(47,710,3122,3468,3770))]
#trydat.SLA <- rtry_exclude(trydata, DataID %in% c(54), baseOn = ObservtionID)

trySLA <- rtry_select_row(input = trydata,getAncillary = TRUE, rmDuplicates = TRUE, TraitID %in% c(3115,3116,3117,3086,47,3122))
#View(rtry_explore(trySLA, DataID, DataName, TraitID, TraitName, sortBy = TraitID))

# get rid of min and max LDMC (and 'NIRS based' whatever that means)
trySLA <- trySLA %>% filter(!DataID %in% c(4027, 4028,4215,7218,7219) ) # removes 1186 values

# remove values with error risk >3 (33356 total rows of uncertain traits that are 3sd >spp/genus/fam mean)
#View(rtry_explore(trySLA, TraitID, TraitName, sortBy = TraitID))

trySLA <- trySLA %>% filter(is.na(ErrorRisk) | ErrorRisk<3)
#View(rtry_explore(tmp, TraitID, TraitName, sortBy = TraitID))

# recalculate LDMC for trait 54 (LWC= 1-LDMC)
#trySLA$StdValue[which(trySLA$DataID==54)] # these don't look any different than real LDMCs...
#plot(OrigValueStr~StdValue, trySLA[which(trySLA$DataID==54),], ylim=c(0,1))
  #Aaaand they've definitely been fixed already. Both for units and 1-LWC


# pivot wider
num_SLA <- rtry_select_row(trySLA, complete.cases(TraitID) & complete.cases(StdValue))
num_SLA <- rtry_select_col(num_SLA, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)

# ancillary data to keep
SLA_lat <- rtry_select_anc(trySLA, 59)
SLA_lon <- rtry_select_anc(trySLA, 60)
SLA_biom <- rtry_select_anc(trySLA, 193)
SLA_devstatus <- rtry_select_anc(trySLA, 413)
SLA_exp <- rtry_select_anc(trySLA, 327)
# add it into the df
num_SLA <- rtry_join_left(num_SLA, SLA_lat, baseOn = ObservationID)
num_SLA <- rtry_join_left(num_SLA, SLA_lon, baseOn = ObservationID)
num_SLA <- rtry_join_left(num_SLA, SLA_biom, baseOn = ObservationID)
num_SLA <- rtry_join_left(num_SLA, SLA_devstatus, baseOn = ObservationID)
num_SLA <- rtry_join_left(num_SLA, SLA_exp, baseOn = ObservationID)

#pivot wider
SLA_dirty <- rtry_trans_wider(num_SLA, names_from=c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn=list(StdValue = mean))

# rename LDMC and SWC
colnames(SLA_dirty)[grep("47", colnames(SLA_dirty))] <- "LDMC"
colnames(SLA_dirty)[grep("3122", colnames(SLA_dirty))] <- "SWC"
#SLA_dirt <- SLA_dirty %>% select(-grep("47", colnames(SLA_dirty))) # clean out the LDMC that stayed around
SLA_dirty$SLA <- apply(SLA_dirty[,grep("SLA", colnames(SLA_dirty))], MARGIN = 1,FUN = mean, na.rm=T)
colnames(SLA_dirty)[6:8] <- c("Biome","Dev_Status","Treatment")

# verify that SWC and LDMC are the same for all the 11347 obs that have both
#swc <- (1/SLA_dirty$LDMC)-1
#plot(swc[which(SLA_dirty$SWC>0)]~SLA_dirty$SWC[which(SLA_dirty$SWC>0)], xlab="SWC from TRY", ylab="SWC from LDMC")
  # oh sweet baby jesus, those are not the same at all.

# remove non-wild grown things

trySLAexp <- rtry_select_row(trySLA, DataID == 327)
trySLAexptbl <- rtry_explore(trySLAexp, DataID, DataName, OriglName, OrigValueStr)

badconditions <- c("C","G","Herbarium","FACE",
                   trySLAexptbl$OrigValueStr[grep("chamber",trySLAexptbl$OrigValueStr)],
                   trySLAexptbl$OrigValueStr[grep("Chamber",trySLAexptbl$OrigValueStr)],
                   trySLAexptbl$OrigValueStr[grep("house",trySLAexptbl$OrigValueStr)], 
                   trySLAexptbl$OrigValueStr[grep("House",trySLAexptbl$OrigValueStr)], 
                   trySLAexptbl$OrigValueStr[grep("pot",trySLAexptbl$OrigValueStr)])
badconditions <- c("Herbarium","FACE") # Doesn't actually save any spp in the overall analysis
SLA <- SLA_dirty %>% filter(!Treatment %in% badconditions) %>% select(-grep("[0-9]", colnames(SLA_dirty)))

 # calcualte SWC from LDMC
SLA$SWC_from_LDMC <- (1/SLA$LDMC) - 1

hist(SLA$SWC_from_LDMC)
hist(SLA$SWC)

length(which(SLA$SWC==SLA$SWC_from_LDMC))
length(which(round(SLA$SWC,1)==round(SLA$SWC_from_LDMC,1)))
experiment <- which(SLA$SWC>0 & SLA$SWC_from_LDMC>0 & round(SLA$SWC_from_LDMC,1)> round(SLA$SWC,1)+.3)

plot(SWC_from_LDMC~SWC, SLA[which(SLA$SWC>0),])
 # oh sweet baby jesus, those are not the same at all.
points(SWC_from_LDMC~SWC, SLA[which(round(SLA$SWC,1)==round(SLA$SWC_from_LDMC,1)),], col="blue")
points(SWC_from_LDMC~SWC, SLA[experiment,], col="green")
plot(SWC_from_LDMC~SWC, SLA[which(SLA$SWC>0),], log="xy")
# doesn't work to exclude for some god forsaken reason
# tried to remove non-field grown measurements

                   

#trySLA.cln <- rtry_exclude(trySLA, (DataID %in% 327) & (OrigValueStr %in% badconditions), baseOn = ObservtionID)

```



SWC = 1/LDMC - 1

But in TRY there are 11347 with both LDMC and SWC, and they don't look the same.

165 are exactly the same
5348 are the same with rounding to 0.1 -> 4951 once removed 3sigma error risk
5999 are quite different. -> 5692 once 3sig removed. That's still over half.

Thoughts: 
- Could be that some LDMC (the scaling high and low ones) are Min and Max values that skewed the average of an obs so they ARENT the exact same. *tested and not it
- all the things on the off-diagonal in log-log space must be... different traits? not different units? *3sig error risk removes the off diagonal by an large
- DataID 54 is maybe not real LDMC? OH! it's defo (1-LDMC). maybe that's part of it? (53k values!) *already corrected

What remains are largely 3 diagonals, 1:1,LDMC 1:1.5ish SWC, smattering of v-low values for SWC
- 

It's possible that the weird highest diagonal is actually 1-LDMC. But the middle one is definitley something about definitions and is not 1-LDMC...



```{r average TRY to spp, include=F}
SLAspp <- SLA %>% group_by(AccSpeciesID, AccSpeciesName) %>% summarise(
  nSLA = n(), SLA=mean(SLA, na.rm=T),
  nLDMC = length(which(LDMC>0)), LDMC = mean(LDMC, na.rm=T),
  nSWC = length(which(SWC>0)), SWC = mean(SWC, na.rm=T))
  # 18,523 species with SLA data, 17892 after cleaning (18078 with greenhouse/growth chamber)
  # 7797 spp with LDMC (7822 with GH/GC)
  # 58 spp with SWC  (58 w/ GH/GC)
SLAspp$SWC_from_LDMC <- (1/SLAspp$LDMC)-1

write.csv(SLAspp, "TRY_SpeciesAverages_20230719.csv")
```