---
title: "Data_Analysis"
author: "LDL Anderegg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Point 1) LMA and SWC aren't highly correlated, and LMA is more variable than SWC. Therefore, LMA drives most of the variation in LWA.


## Saturated Water Content: How much water is there per dry mass ('SWC') vs m2 leaf area ('SWA')
``` {r Saturated water per mass vs area, echo=F, warnings=F }
par(mfrow=c(2,1),mar=c(8,8,2,1))
boxplot(SWC~Biome, fullspp, las=2
        , xlab="", ylab=expression(paste("Sat g water per g dry mass")))
mtext(text=paste("# Species=",length(which(fullspp$SWC>0))), side=3,line = 0, adj = 0)

boxplot(SWA~Biome, fullspp, las=2
        , xlab="", ylab=expression(paste("Sat g water per ",m^2," leaf area")))
mtext(text=paste("# Species=",length(which(fullspp$SWA>0))), side=3,line = 0, adj = 0)

```
*Takehome*: LMA really really matters. \
-Crops are the wettest and Med/Conifers driest saturated water content (SWC) per unit dry mass\
-But conifers have the highest saturated water content per unit leaf area, and crops and Med move towards the mean. Temperate plants have the least water per unit leaf area @ saturation.\



``` {r log-log SWC and LMA plot}
mypal <- brewer.pal(n=8, "Dark2")
palette(mypal)
par(mar=c(6,6,2,1))
plot(LMA~SWC, fullpvs, pch=16, col="grey", log="xy", ylab="LMA", xlab="Sat g Water per g dry mass\nMass Basis")
palette(mypal)
points(LMA~SWC, fullspp, pch=1, col=factor(Gymno.Angio))
points(LMA~SWC, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=0.2, y=500, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)
```

\
Indeed, there's a vague decrease in LMA with increasing SWC among Angiosperm species, but really no strong relationship between the two. So the SWC~SWA relationship is going to be hard to predict from LMA or SWC.\
\
\


### log-log relationships

``` {r continued SWC on mass vs area basis}


# ggplot(fullspp, aes(x=SWC,y=SWA, col=Biome)) +  xlab("Sat g water per g dry mass\nMass Basis") + ylab("Sat g water per m2\nArea Basis") + geom_point(data=fullpvs) + geom_point(col="#00000066") + ggtitle("All Obs, raw") + facet_wrap(~Gymno.Angio)

# ggplot(fullspp, aes(x=log(SWC, base=10),y=log(SWA,base=10), col=Biome)) +  xlab("log10 Sat g water per g dry mass\nMass Basis") + ylab("log10 Sat g water per m2\nArea Basis") + geom_smooth(method="lm", se=F) + #geom_point(data=fullpvs) + 
#   geom_point() + ggtitle("Spp means, log-log")

# NOTE: the normal geom_smooth really shows different scaling lines than typeII regression.
#ggplot(pvs, aes(x=log(LMA),y=log(SWA), col=Biome)) + geom_point() + xlab("log(LMA)") + ylab("log(g water per m2 leaf)")

#ggplot(pvs, aes(x=LMA,y=SWC, col=Biome)) + geom_point() + xlab("LMA") + ylab("g water per m2 leaf")

# \
# 
# - There is generally a positive scaling relationship between mass- versus area-based saturated water content, but there are strong intercept differences between biomes and possibly some scaling slope differences\
# - There's definitely a much steeper slope in SWC on an Area vs Mass basis across gymnosperms than angiosperms. scaling slope = LMA so makes sense that gymnos have higher LMA (if there's no change in LMA/correlation with SWC)

```

``` {r log-log plot}
mypal <- brewer.pal(n=8, "Dark2")
palette(mypal)
par(mar=c(6,6,2,1))
plot(SWA~SWC, fullpvs, pch=16, col="grey", log="xy", ylab="SWA\n(gH2O m-2 Area)", xlab="SWC\n(gH2O g-1 dry mass)")
palette(mypal)
points(SWA~SWC, fullspp, pch=1, col=factor(Gymno.Angio))
points(SWA~SWC, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=0.15, y=300, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=1, bty="n", xpd=NA)
# SMA regression for gymnos
gymnomod <- lmodel2::lmodel2(lm(log(SWA, 10)~I(log(SWC,10)), fullspp[which(fullspp$Gymno.Angio=="Gymnosperm"),]))
abline(a=gymnomod$regression.results$Intercept[3], b=gymnomod$regression.results$Slope[3], col=2)

angiomod <- lmodel2::lmodel2(lm(log(SWA, 10)~I(log(SWC,10)), fullspp[which(fullspp$Gymno.Angio=="Angiosperm"),]))
abline(a=angiomod$regression.results$Intercept[3], b=angiomod$regression.results$Slope[3], col=1)


redwoodmod <- lmodel2::lmodel2(lm(log(SWA, 10)~I(log(SWC,10)), fullpvs[which(fullpvs$Species=="Sequoiadendron giganteum"),]))
abline(a=redwoodmod$regression.results$Intercept[3], b=redwoodmod$regression.results$Slope[3], col="gray")

pinemod <- lmodel2::lmodel2(lm(log(SWA, 10)~I(log(SWC,10)), fullpvs[which(fullpvs$Species=="Pinus albicaulis"),]))
abline(a=pinemod$regression.results$Intercept[3], b=pinemod$regression.results$Slope[3], col="gray")

# oakmod <- lmodel2::lmodel2(lm(log(SWA, 10)~I(log(SWC,10)), fullpvs[which(fullpvs$Species=="Quercus garryana"),]))
# abline(a=oakmod$regression.results$Intercept[3], b=oakmod$regression.results$Slope[3], col="gray")


```
- among gymnsoperm species, there's a steeper scaling relationship between SWA and SWC 
- however, within a gymnosperm species (see the redwood cluster of gray points) the scaling slope looks similar to the angiosperm slope but with a massive offset in intercept\
\





## Role of SWC and LMA in determining Total Water Loss
The total amount of water *per unit leaf area* (total water loss per m2, or TWL) that can be lost from a leaf between saturation and TLP is a product of the RWC at TLP (RWCtlp), the Saturated Water Content (SWC) and Leaf Mass per Area (LMA).

``` {r visualizing families}
mypal <- brewer.pal(n=8, "Dark2")
par(mar=c(6,6,2,1))
plot(water_loss.m2~SWC, fullpvs, pch=16, col="grey", log="xy", ylab="TWL, g water m-2\n(sat to TLP)")
palette(mypal)
points(water_loss.m2~SWC, fullspp, pch=1, col=factor(Gymno.Angio))
points(water_loss.m2~SWC, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=0.2, y=500, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)

par(mar=c(6,6,2,1))
plot(water_loss.m2~LMA, fullpvs, pch=16, col="grey", log="xy", ylab="TWL g water m-2\n(sat to TLP)")
palette(mypal)
points(water_loss.m2~LMA, fullspp, pch=1, col=factor(Gymno.Angio))
points(water_loss.m2~LMA, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=20, y=500, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)

par(mar=c(6,6,2,1))
plot(water_loss.m2~totRWCtlp, fullpvs, pch=16, col="grey", log="xy", ylab="TWL g water m-2\n(sat to TLP)")
palette(mypal)
points(water_loss.m2~totRWCtlp, fullspp, pch=1, col=factor(Gymno.Angio))
points(water_loss.m2~totRWCtlp, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=0.2, y=1100, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)


par(mar=c(6,6,2,1))
plot(SWA~LMA, fullpvs, pch=16, col="grey", log="xy", ylab="SWA \n(g H2O m-2)")
palette(mypal)
points(SWA~LMA, fullspp, pch=1, col=factor(Gymno.Angio))
points(SWA~LMA, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend("bottom", legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)
```

#### Gymno-Angio divide in what drives TWL:
-All the variation in TWL in *Gymnosperms* is due to LMA, and confers have massive LMA so they have high TWL. Also strongly related to totRWCtlp within sequoia and somewhat across species\

-The variation in *Angiosperms* isn't strongly driven by LMA, but jointly determined by SWC and totRWCtlp.\
``` {r SWC~RWCtlp}
plot(SWC~totRWCtlp, fullpvs, pch=16, col="grey", log="xy")
palette(mypal)
points(SWC~totRWCtlp, fullspp, pch=1, col=factor(Gymno.Angio))
points(SWC~totRWCtlp, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=0.2, y=1100, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)

## joint control of SWC and totRWCtlp
```
\
Interestingly, there's essentially no relationship\
So in Angiosperms the effects of the two are largely independent:

``` {r effects of totRWC and SWC on TWL}
ggplot(fullspp[which(fullspp$Gymno.Angio=="Angiosperm"),], aes(x=log(SWC,10), y=log(water_loss.m2, 10), col=totRWCtlp)) + geom_point() + geom_smooth(data=fullspp[which(fullspp$Gymno.Angio=="Angiosperm" & fullspp$totRWCtlp<80),], method="lm", se=F, col="black") + geom_smooth(data=fullspp[which(fullspp$Gymno.Angio=="Angiosperm" & fullspp$totRWCtlp>80),], method="lm", se=F) + ylab("log10 TWL, g water m-2")
```

(black line in last plot is spp with totRWCtlp < 80% and blue is spp with totRWCtl>80%)\


## How you accomplish high TWL
TWL probably has a stronger methodological implication (for remote sensing) than physiological implication. But  it might be worth exploring how species with high TWL accomplish this large water loss. \
Is it do to high capacitance (CFT, g water per m2 leaf area per MPa water potential change)?\
Or to low RWC @ TLP (i.e. the ability to desiccate your leaves)?

First, are capacitance (CFT) and RWCtlp related? (i.e. could it be both?)
```{r How you accomplish high TWL}
# Megan's point about achieving high water loss either by having high capacitance or low totRWCtlp
par(mar=c(6,6,2,1))
plot(CFT.g.m2~totRWCtlp, fullpvs, pch=16, col="grey", log="xy")
palette(mypal)
points(CFT.g.m2~totRWCtlp, fullspp, pch=1, col=factor(Gymno.Angio))
points(CFT.g.m2~totRWCtlp, fullfam, pch=16, col=factor(Gymno.Angio))
legend('topleft', legend=c("obs", "species","family"), pch=c(16,1,16), col=c("grey","black","black"), bty="n", cex=.9)
legend(x=0.2, y=1100, legend=c("Gymno","Angio"), pch=16, col=mypal[c(2,1)], ncol=2, bty="n", xpd=NA)

```
\
They are somewhat negatively related, so high TWL is a bit of both. But they're not as related as you might expect.

Looking at the log-log scaling:
``` {r log-log scaling of CFT and RWCtlp}
ggplot(pvs[which(pvs$water_loss.m2>0),], aes(y=CFT.g.m2_mycalc, x=totRWCtlp, col=log(water_loss.m2,base=10))) + geom_point()
#ggplot(fullpvs[which(fullpvs$water_loss.m2>0),], aes(y=water_loss.m2, x=totRWCtlp, col=CFT.g.m2)) + geom_point()
```
\
### Takehome: \
It looks to me like CFT (capacitance) is actually more important than totRWCtlp for driving high TWL on a leaf area basis. 


## Decomposing Variance: 
#### How much and why do all of these variables vary globally?

*1) The absolute x-fold variation* \
(calculated here as (max - min)/min...is this the right way to calculate this?):
``` {r calcualting xfold variation}
# make some functions for quantifying variation
xfold.full <- function(data){
  # data (max-min)/min
  round((max(data, na.rm=T)- min(data, na.rm=T))/min(data,na.rm=T),2)
  }
xfold.quant <- function(data, quant.range=c(.25,.75)){
  # (high quantile - low quantile) / low quantile
  round((quantile(x=data, probs=quant.range[2], na.rm=T) - quantile(x=data, probs=quant.range[1], na.rm=T))/quantile(x=data, probs=quant.range[1], na.rm=T),2)
}

IQR <- function(data, quant.range=c(.25,.75)){
  # (high quantile - low quantile) - Interquartile Range
  round((quantile(x=data, probs=quant.range[2], na.rm=T) - quantile(x=data, probs=quant.range[1], na.rm=T)),2)
}

# cofeeficient of variation
CV <- function(data){
  sd(data, na.rm=T)/mean(data, na.rm=T)
}

# make a dataframe for storing variation quantification info
varmags <- data.frame(variabele=c("TLP","RWCtlp","SWC","SWA","LMA","C[T]","TWL"), xfold.raw =rep(NA, 7), xfold.90 =rep(NA, 7), xfold.50=rep(NA, 7), IQR =rep(NA, 7), sd =rep(NA, 7), CV = rep(NA, 7))

varmags$xfold.raw <- c(xfold.full(-1*fullpvs$TLP),
                       xfold.full(fullpvs$totRWCtlp),
                       xfold.full(fullpvs$SWC),
                       xfold.full(fullpvs$SWA),
                       xfold.full(fullpvs$LMA),
                       xfold.full(fullpvs$CFT.g.m2),
                       xfold.full(fullpvs$water_loss.m2))

varmags$xfold.90 <- c(xfold.quant(-1*fullpvs$TLP, quant.range = c(0.05,.95)),
                       xfold.quant(fullpvs$totRWCtlp, quant.range = c(0.05,.95)),
                       xfold.quant(fullpvs$SWC, quant.range = c(0.05,.95)),
                       xfold.quant(fullpvs$SWA, quant.range = c(0.05,.95)),
                       xfold.quant(fullpvs$LMA, quant.range = c(0.05,.95)),
                       xfold.quant(fullpvs$CFT.g.m2, quant.range = c(0.05,.95)),
                       xfold.quant(fullpvs$water_loss.m2, quant.range = c(0.05,.95)))

varmags$xfold.50 <- c(xfold.quant(-1*fullpvs$TLP),
                       xfold.quant(fullpvs$totRWCtlp),
                       xfold.quant(fullpvs$SWC),
                       xfold.quant(fullpvs$SWA),
                       xfold.quant(fullpvs$LMA),
                       xfold.quant(fullpvs$CFT.g.m2),
                       xfold.quant(fullpvs$water_loss.m2))

varmags$IQR <- c(IQR(-1*fullpvs$TLP),
                       IQR(fullpvs$totRWCtlp),
                       IQR(fullpvs$SWC),
                       IQR(fullpvs$SWA),
                       IQR(fullpvs$LMA),
                       IQR(fullpvs$CFT.g.m2),
                       IQR(fullpvs$water_loss.m2))

varmags$sd <- c(sd(-1*fullpvs$TLP, na.rm=T),
                       sd(fullpvs$totRWCtlp, na.rm=T),
                       sd(fullpvs$SWC, na.rm=T),
                       sd(fullpvs$SWA, na.rm=T),
                       sd(fullpvs$LMA, na.rm=T),
                       sd(fullpvs$CFT.g.m2, na.rm=T),
                       sd(fullpvs$water_loss.m2, na.rm=T))


varmags$CV <- c(CV(-1*fullpvs$TLP),
                       CV(fullpvs$totRWCtlp),
                       CV(fullpvs$SWC),
                       CV(fullpvs$SWA),
                       CV(fullpvs$LMA),
                       CV(fullpvs$CFT.g.m2),
                       CV(fullpvs$water_loss.m2))

# reorder based on my conceptual introduction
varmags <- varmags[c(1,2,5,3,4,6,7),]

print(varmags)
```
``` {r xfold plot}
quartz(width=3, height=3)
par(mgp=c(2.5,1,0))
bp <- barplot(height=varmags$xfold.90[c(1,2,4,3,5,6,7)], names.arg = varmags$variabele[c(1,2,4,3,5,6,7)], las=2, ylab="X-fold variation", col="#66666666")
#barplot(height=varmags$CV*11/0.77, names.arg=NA, add=T, col="#00006666", yaxt="n")

```

*TWL, CFT, LMA and SWA are WAY more variable than TLP and RWCtlp* 
Depending on what metric you look at (xfold.raw is x-fold variation in all data, so prone to outliers, xfold-90 is just the variation from the 5ht -95th quantiles, CV assumes a normal distribution which many of these aren't), either TWL or LMA are the most variable metrics.

### Taxonomic Variance Decomposition
If we decompose the variance of various elements by the taxonomic scale, we see that all of these attributes show way more variation within-species than most leaf traits (e.g. LES traits)


``` {r variance decomp, include=F}


require(lme4)
require(lmerTest)
require(MuMIn)

# # or fitting with a Bayesian model
# require(rstanarm)
# require(ggplot2)
# require(bayesplot)
# theme_set(bayesplot::theme_default())
# Raw Traits
TLPmod <- lmer(TLP~1 + (1|Family/Genus/Species), fullpvs)
RWCtlpmod <- lmer(totRWCtlp~1 + (1|Family/Genus/Species), fullpvs)
LMAmod <- lmer(LMA~1 + (1|Family/Genus/Species), fullpvs)
SWCmod <- lmer(SWC~1 + (1|Family/Genus/Species), fullpvs)
gsatmod <- lmer(SWA~1 + (1|Family/Genus/Species), fullpvs)
CFTmod <- lmer(CFT.g.m2~1 + (1|Family/Genus/Species), fullpvs)
WLGmod <- lmer(water_loss.g~1 + (1|Family/Genus/Species), fullpvs)
WLMmod <- lmer(log(water_loss.m2,10)~1 + (1|Family/Genus/Species), fullpvs)
# WLMmod <- stan_lmer(water_loss.m2~1 + (1|Family/Genus/Species), fullpvs) # yup, the Family sigma is low
# SWCmod <- stan_lmer(log(SWC,10)~1 + (1|Family/Genus/Species), fullpvs)
# launch_shinystan(SWCmod)
  #well, it actually DOES look like the Family sigma gets estimated at ~0 in both ML and Bayes land for log(SWC)
  # I don't know why this varies so much from RAW to Logged for SWC...

TLPvar<- data.frame(VarCorr(TLPmod))
RWCtlpvar<- data.frame(VarCorr(RWCtlpmod))
LMAvar <- data.frame(VarCorr(LMAmod))
SWCvar<- data.frame(VarCorr(SWCmod))
gsatvar<- data.frame(VarCorr(gsatmod))
CFTvar<- data.frame(VarCorr(CFTmod))
WLGvar<- data.frame(VarCorr(WLGmod))
WLMvar <- data.frame(VarCorr(WLMmod))


traitvars <- data.frame(TLPvar[,4], RWCtlpvar[,4], LMAvar[,4],SWCvar[,4], gsatvar[,4], CFTvar[,4], WLGvar[,4], WLMvar[,4])
colnames(traitvars) <- c("TLP","RWCtlp","LMA", "SWC","SWA", "CFT", "WaterLoss_g", "TWL")
rownames(traitvars) <- c("BtwSpecies", "BtwGenera", "BtwFamilies", "WtinSpecies")

traitvars <- traitvars[c("BtwFamilies","BtwGenera","BtwSpecies","WtinSpecies"),]

traitvars_scaled <- traitvars
for(i in 1:ncol(traitvars)){
  traitvars_scaled[,i] <- traitvars[,i]/sum(traitvars[,i])
}


```
red = btw families,
dark blue = btw genera w/in family,
blue = btw spp w/in genus,
light blue = within species
``` {r var decomp plot}
#quartz(width=5, height=4)
cols <- brewer.pal(11, "RdBu")[c(1,11,9,7)]
# cols <- "#B2182B" "#053061" "#F7F7F7" "#4393C3"
#"#67001F" "#053061" "#4393C3" "#D1E5F0"
barplot(as.matrix(traitvars_scaled),beside=F,legend.text = F,xpd = T, names.arg = c("TLP","RWCtlp","LMA", "SWC g-1","SWC m-2","CFT","TWL g-1", "TWL m-2"), las=2,args.legend = list(x=4, y=1.3, ncol=2), col = paste0(cols,"CC"), ylab="Proportion of total Variance", xlab="")
legend(xpd=T, x = 0, y=2.3, legend=c("W/Spp", "Btw Spp", "Btw Genera", "Btw Families"), fill=paste0(cols,"CC")[c(4,3,2,1)], ncol=2, bty="n",  cex=1.2)

#  hist(log(-1*fullpvs$TLP)) # prob log needed for TLP
#  hist(fullpvs$totRWCtlp) # log doesn't help
#  hist(log(fullpvs$SWC)) # prob log needed for SWC
#  hist(fullpvs$SWA) # maybe needed for SWA (normalizes the non-Sequoias)
#  hist(log(fullpvs$CFT.g.m2)) # maybe needed for CFT (kinda 0 truncated)?
#  hist(fullpvs$)
#  # logged Traits
# TLPmod <- lmer(log(TLP*-1,10)~1 + (1|Family/Genus/Species), fullpvs)
# SWCmod <- lmer(SWC~1 + (1|Family/Genus/Species), fullpvs)
# CFTmod <- lmer(log(CFT.g.m2,10)~1 + (1|Family/Genus/Species), fullpvs)
# WLmod <- lmer(log(water_loss.m2,10)~1 + (1|Family/Genus/Species), fullpvs)
# # TLP and SWC defo need it. CFT prob not and water_loss maybe not
# 
# SWCvar<- data.frame(VarCorr(SWCmod))
# TLPvar<- data.frame(VarCorr(TLPmod))
# CFTvar<- data.frame(VarCorr(CFTmod))
# WLvar<- data.frame(VarCorr(WLmod))
# 
# traitvars <- data.frame(TLPvar[,4], SWCvar[,4], CFTvar[,4], WLvar[,4])
# colnames(traitvars) <- c("TLP", "SWC", "CFT", "WaterLoss")
# rownames(traitvars) <- c("BtwSpecies", "BtwGenera", "BtwFamilies", "WtinSpecies")
# 
# traitvars <- traitvars[c("BtwFamilies","BtwGenera","BtwSpecies","WtinSpecies"),]
# 
# traitvars_scaled <- traitvars  
# for(i in 1:ncol(traitvars)){
#   traitvars_scaled[,i] <- traitvars[,i]/sum(traitvars[,i])
# }

# calculate x-fold variation as well as CV for contextualizing how variable traits are




```

``` {r var decomp for rarified dataset}
# need to rarify Sequoiadendron giganteum because too damn many of them.
# maybe also Pinus albicaulis?

rarepvs <- fullpvs[which(fullpvs$Species!="Sequoiadendron giganteum" & fullpvs$Species != "Pinus albicaulis"),]
SEGI <- fullpvs[sample(which(fullpvs$Species=="Sequoiadendron giganteum"),size =34),]
PIAB <- fullpvs[which(fullpvs$Species=="Pinus albicaulis" & fullpvs$SWC>0),] 
# only 34 PIABs have actual SWC and LMA values anyway

rarepvs <- rbind(rarepvs, SEGI, PIAB)


# # or fitting with a Bayesian model
# require(rstanarm)
# require(ggplot2)
# require(bayesplot)
# theme_set(bayesplot::theme_default())
# Raw Traits
TLPmod <- lmer(TLP~1 + (1|Family/Genus/Species), rarepvs)
RWCtlpmod <- lmer(totRWCtlp~1 + (1|Family/Genus/Species), rarepvs)
LMAmod <- lmer(LMA~1 + (1|Family/Genus/Species), rarepvs)
#LMAmod <- lmer(log(LMA)~1 + (1|Family/Genus/Species), rarepvs)
SWCmod <- lmer(SWC~1 + (1|Family/Genus/Species), rarepvs[!rownames(rarepvs) %in% c(931, 768),])
gsatmod <- lmer(log(SWA, 10)~1 + (1|Family/Genus/Species), rarepvs[!rownames(rarepvs) %in% c(1014, 622),])
CFTmod <- lmer(log(CFT.g.m2,10)~1 + (1|Family/Genus/Species), rarepvs)
WLGmod <- lmer(water_loss.g ~1 + (1|Family/Genus/Species), rarepvs[!rownames(rarepvs) %in% c(768, 931),]) # raw outliers [!rownames(rarepvs) %in% c(768, 931),], 1014 logged
WLMmod <- lmer(log(water_loss.m2,10)~1 + (1|Family/Genus/Species), rarepvs)
# WLMmod <- stan_lmer(water_loss.m2~1 + (1|Family/Genus/Species), rarepvs) # yup, the Family sigma is low
# SWCmod <- stan_lmer(log(SWC,10)~1 + (1|Family/Genus/Species), rarepvs)
# launch_shinystan(SWCmod)
  #well, it actually DOES look like the Family sigma gets estimated at ~0 in both ML and Bayes land for log(SWC)
  # I don't know why this varies so much from RAW to Logged for SWC...

# Log transforms?
# qqp(resid(TLPmod)) # maybe better, but trades high for low resids
# qqp(resid(RWCtlpmod)) #almost no effect
# qqp(resid(LMAmod)) # does kinda improves. Logging increases within/genus var
# qqp(resid(SWCmod)) # ok if you remove outliers. doesn't need it.
# qqp(resid(gsatmod)) # looks substantially better logged, and hist looks better
# qqp(resid(CFTmod)) # looks quite a bit better, as does hist, shrinks w/in Fam and a bit fam and w/in spp, increases w/in genus
# qqp(resid(WLGmod)) # slight improvement with logging, hist looks a bit better. TOTALLY changes interpretation from mostly Fam to mostly not.... YIKES
# qqp(resid(WLMmod)) # logging expands w/in genus, but doesn't change Fam

TLPvar<- data.frame(VarCorr(TLPmod))
RWCtlpvar<- data.frame(VarCorr(RWCtlpmod))
LMAvar <- data.frame(VarCorr(LMAmod))
SWCvar<- data.frame(VarCorr(SWCmod))
gsatvar<- data.frame(VarCorr(gsatmod))
CFTvar<- data.frame(VarCorr(CFTmod))
WLGvar<- data.frame(VarCorr(WLGmod))
WLMvar <- data.frame(VarCorr(WLMmod))


traitvars_rar <- data.frame(TLPvar[,4], RWCtlpvar[,4], LMAvar[,4],SWCvar[,4], gsatvar[,4], CFTvar[,4], WLGvar[,4], WLMvar[,4])
colnames(traitvars_rar) <- c("TLP","RWCtlp","LMA", "SWC","SWA", "CFT", "WaterLoss_g", "TWL")
rownames(traitvars_rar) <- c("BtwSpecies", "BtwGenera", "BtwFamilies", "WtinSpecies")

traitvars_rar <- traitvars_rar[c("BtwFamilies","BtwGenera","BtwSpecies","WtinSpecies"),]

traitvars_rar_scaled <- traitvars_rar
for(i in 1:ncol(traitvars_rar)){
  traitvars_rar_scaled[,i] <- traitvars_rar[,i]/sum(traitvars_rar[,i])
}


#quartz(width=5, height=4)
# quartz(width=5, height=3.5)
# par(mgp=c(3,.7,0), cex.lab=1.3, cex.axis=1.1, mfrow=c(1,1), mar=c(6,2,3,6), oma=c(0,3,0,0))
# 
# mypal <- brewer.pal(n=9, "Set1")
# colchoices <- c(1,2,4,3,6)
# cols <-rev(c(mypal[colchoices[c(1,2,3)]],"#CCCCCC"))#brewer.pal(11, "RdBu")[c(11,9,1, 6)]
# # "#CCCCCC" "#984EA3" "#377EB8" "#E41A1C"
# #cols <- brewer.pal(11, "RdBu")[c(1,11,9,7)]
# # cols <- "#B2182B" "#053061" "#F7F7F7" "#4393C3"
# #"#67001F" "#053061" "#4393C3" "#D1E5F0"
# barplot(as.matrix(traitvars_rar_scaled[,-7]),beside=F,legend.text = F,xpd = T, names.arg = c("TLP","RWCtlp","LMA", "SWC","SWA","C[T]", "TWL"), las=2,args.legend = list(x=4, y=1.3, ncol=2), col = paste0(cols,"99"), ylab="Proportion of total Variance", xlab="")
# mtext("Proportion of Total Variance", side=2, line=2.8)
# 
# #legend(xpd=T, x = 0, y=2.3, legend=c("W/Spp", "Btw Spp", "Btw Genera", "Btw Families"), fill=paste0(cols,"CC")[c(4,3,2,1)], ncol=2, bty="n",  cex=1.2)
# legend(xpd=NA, x = 8.5, y=0.7, legend=rev(c("btw Fams","btw Gen","btw Spp","w/in Spp")), fill=rev(paste0(cols,"99")), ncol=1, bty="n",  cex=1)
# 

```

\
-TLP: Turgor Loss Point (MPa)\
- SWC/TWL g-1 = Sat Water Content/Total Water Loss per unit leaf mass\
- SWC/TWL m-2 = per unit leaf area.\
*note: these may be sensitive to the large Sequoiadendron dataset, need to double check that. 
\
\
### Look at some of the same attributes per biome and leaf habit
-removed Picea mariana because massive high outlier
```{r prelim analysis}
par(mar=c(8,5,2,1))
boxplot(water_loss.m2~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),], las=2, ylab="TWL, g water m-2\nfull sat to tlp", xlab="")
mtext(text=paste("# Species=",length(unique(fullspp$Species[which(fullspp$water_loss.m2>0)]))), side=3,line = 0, adj = 0)
mtext(text=paste("Biome R2:",round(summary(lm(water_loss.m2~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-1,adj=0)
mtext(text=paste("Ever.Decid R2:",round(summary(lm(water_loss.m2~Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-2,adj=0)

boxplot(SWC~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),], las=2, ylab="SWC\ng water g-1 dry mass", xlab="")
mtext(text=paste("# Species=",length(unique(fullspp$Species[which(fullspp$water_loss.m2>0)]))), side=3,line = 0, adj = 0)
mtext(text=paste("Biome R2:",round(summary(lm(SWC~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-1,adj=0)
mtext(text=paste("Ever.Decid R2:",round(summary(lm(SWC~Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-2,adj=0)

boxplot(CFT.g.m2~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),], las=2, ylab=expression(paste("CFT, g water lost ", MPa^-1, m^-2)), xlab="")
mtext(text=paste("# Species=",length(unique(fullspp$Species[which(fullspp$CFT.g.m2>0)]))), side=3,line = 0, adj = 0)
mtext(text=paste("Biome R2:",round(summary(lm(CFT.g.m2~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-1,adj=0)
mtext(text=paste("Ever.Decid R2:",round(summary(lm(CFT.g.m2~Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-2,adj=0)

boxplot(TLP~Biome, fullspp, las=2, ylab="TLP (MPa)", xlab="")
mtext(text=paste("# Species=",length(unique(fullspp$Species[which(fullspp$TLP>0)]))), side=3,line = 0, adj = 0)
mtext(text=paste("Biome R2:",round(summary(lm(TLP~Biome, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-1,adj=0)
mtext(text=paste("Ever.Decid R2:",round(summary(lm(TLP~Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))$r.squared, 2)),line=-2,adj=0)
#boxplot(LMA~Biome, fullspp, las=2, ylab="LMA", xlab="")
#mtext(text=paste("# Species=",length(unique(fullspp$Species[which(fullspp$LMA>0)]))), side=3,line = 0, adj = 0)

```
\
Looks like Biomes are way more differentiable for TLP (and LMA) than actual water being lost...
``` {r more plotzz, include=F}
summary(aov(TLP~Biome + Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))
summary(aov(SWC~Biome + Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))
summary(aov(CFT.g.m2~Biome + Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))
summary(aov(water_loss.m2~Biome + Ever.Decid, fullspp[-which(fullspp$Species=="Picea mariana"),]))

```
\
Ultimately, they've got very different modes of variation. \
- TWL doesn't differ by biome OR leaf habit\
- SWC & CFT differ more by Leaf Habit than Biome\
- TLP differs more by biome

\
Assuming that CFT (slope of the Psi~Water/area relationship) and TWL (potential maximum signal) are the two most important metrics, let's explore other predictors of variation for them.
``` {r other drivers CFT and TWL}
#boxplot(CFT.g.m2~Growth.Form, fullspp)
#boxplot(water_loss.m2~Growth.Form, fullspp)
plot(CFT.g.m2~water_loss.m2, fullspp, col=factor(Growth.Form), log="xy", pch=16, ylab="CFT (g m-2 MPa-1)", xlab="TWL (g m-2)")
legend("topleft",c("fern","herb","woody"), pch=16, col=c(1,2,3))
```
Herbaceous and woody plants have fairly similar TWL (woody gets lower), but herbs generally have higher CFT.

``` {r other drivers CFT and TWL2}
boxplot(CFT.g.m2~Ever.Decid, fullspp)
boxplot(water_loss.m2~Ever.Decid, fullspp)
ggplot(fullspp, aes(x=log(water_loss.m2, 10), y=log(CFT.g.m2,10), col=Ever.Decid)) + geom_point() + geom_smooth(method="lm")+ theme_classic() + ylab("CFT (g m-2 MPa-1)") + xlab("TWL (g m-2)")
#plot(CFT.g.m2~water_loss.m2, fullspp, col=factor(Ever.Decid), log="xy", pch=16, ylab="CFT (g m-2 MPa-1)", xlab="TWL (g m-2)")
#legend("topleft",c("fern","herb","woody"), pch=16, col=c(1,2,3))
```

``` {r other drivers CFT and TWL3}
boxplot(CFT.g.m2~T.S.L.H, fullspp)
boxplot(water_loss.m2~T.S.L.H, fullspp)
ggplot(fullspp, aes(x=log(water_loss.m2, 10), y=log(CFT.g.m2,10), col=T.S.L.H)) + geom_point() + geom_smooth(method="lm")+ theme_classic() + ylab("CFT (g m-2 MPa-1)") + xlab("TWL (g m-2)")
#plot(CFT.g.m2~water_loss.m2, fullspp, col=factor(Ever.Decid), log="xy", pch=16, ylab="CFT (g m-2 MPa-1)", xlab="TWL (g m-2)")
#legend("topleft",c("fern","herb","woody"), pch=16, col=c(1,2,3))
```



``` {r full models}


cleanrarepvs <- rarepvs[complete.cases(rarepvs %>% select(TLP,CFT.g.m2, water_loss.m2, SWA, Biome, Growth.Form, Ever.Decid, T.S.L.H, Wild.Garden, Life.Stage, Gymno.Angio)),] # removed Growth.Form and Life.Stage, didn't clear any more 

# # check to make sure it's worth fitting everything
# xtabs(~Biome + Growth.Form, cleanrarepvs) # Growth.Form is useless at present, need to combine Coastal with something
# xtabs(~Ever.Decid + T.S.L.H, cleanrarepvs) # both good.
# xtabs(~Wild.Garden + Life.Stage, cleanrarepvs) # pretty good, though only seedlings in pots.
# xtabs(~Species + Life.Stage, cleanrarepvs) 
  # Only 3 spp with more than one life stage
  # c("Arbutus unedo", "Paullinia pinnata", "Vitis vinifera")
boxplot(water_loss.m2~ Life.Stage+Species , cleanrarepvs[which(cleanrarepvs$Species %in% c("Arbutus unedo", "Paullinia pinnata", "Vitis vinifera")),], las=2)
  # inconsistent (higher, lower, same)
boxplot(TLP~ Life.Stage+Species , cleanrarepvs[which(cleanrarepvs$Species %in% c("Arbutus unedo", "Paullinia pinnata", "Vitis vinifera")),], las=2)
  # no difference

  #### DREDGE approach
options(na.action="na.fail")
  # NOTE: I removed Growth.Form (only two non-woody) and Life.Stage (only three species with comparisons)
LWAfullmod <- lmer(SWA~Biome + Ever.Decid + T.S.L.H + Wild.Garden  + Gymno.Angio + (1|Species), cleanrarepvs)
  # 838 obs with all of the above?!? seems actually pretty reasonable...
LWAdredge <- dredge(LWAfullmod)
  # best model is - Grw.Frm, but 2nd best is Grw.Frm # in lmerland it's all
TLPfullmod <- lmer(TLP~Biome + Ever.Decid + T.S.L.H + Wild.Garden  + Gymno.Angio + (1|Species), cleanrarepvs)
  # 838 obs with all of the above?!? seems actually pretty reasonable...
TLPdredge <- dredge(TLPfullmod)
  # best model is Biome + Life.Stage + Wld.Grd (add in Gymno.Angio for 2nd mod)

CFTfullmod <- lmer(CFT.g.m2~Biome + Ever.Decid + T.S.L.H + Wild.Garden  + Gymno.Angio + (1|Species), cleanrarepvs)
  # 838 obs with all of the above?!? seems actually pretty reasonable...
CFTdredge <- dredge(CFTfullmod)
  # best model is Everything

TWLfullmod <- lmer(water_loss.m2~Biome + Ever.Decid + T.S.L.H + Wild.Garden  + Gymno.Angio + (1|Species), cleanrarepvs)
  # 838 obs with all of the above?!? seems actually pretty reasonable...
TWLdredge <- dredge(TWLfullmod)
  # best model is everything

TLPdredge # now Biome + Wld.Grd best, + TSL v close + Gymn.Ango kinda close. Old: just Life.Stage with rarepvs, +Biome+Wld.Grd are next in line
LWAdredge # everything with rare and full
CFTdredge # everything with rare and full
TWLdredge # everythiung with rare and full


options(na.action="na.omit")




library(variancePartition)

cleanrarepvs <- rarepvs[complete.cases(rarepvs %>% select(SWA, TLP, CFT.g.m2, water_loss.m2,Biome, Growth.Form, Ever.Decid, T.S.L.H, Wild.Garden, Life.Stage, Gymno.Angio)),]

# let's just try for LWA before doing all
form.fixed <- ~ Biome+ Ever.Decid + T.S.L.H + Wild.Garden + Gymno.Angio 
form.rand <- ~ (1|Biome) + (1|Ever.Decid) + (1|T.S.L.H)+ (1|Wild.Garden) + (1|Gymno.Angio) + (1|Species)
test.fixed <- fitExtractVarPartModel(exprObj = t(cleanrarepvs %>% select( TLP,SWA, CFT.g.m2, water_loss.m2 )), formula = form.fixed, data= cleanrarepvs)
test.rand <- fitExtractVarPartModel(exprObj = t(cleanrarepvs %>% select(TLP,SWA,  CFT.g.m2, water_loss.m2 )), formula = form.rand, data= cleanrarepvs)
test.fixed <- sortCols(test.fixed)
plotPercentBars(test.fixed)
test.rand <- data.frame(test.rand) %>% select(c(as.character(colnames(test.fixed)), "Species"))
plotPercentBars(test.rand)




par(mgp=c(3,.7,0), cex.lab=1.3, cex.axis=1.1, mfrow=c(1,1), mar=c(6,2,3,6), oma=c(0,3,0,0))
 
mypal <- brewer.pal(n=9, "Set1")
 # "#E41A1C" "#377EB8" "#4DAF4A" "#984EA3" "#FF7F00" "#FFFF33" "#A65628" "#F781BF" "#999999"

 barplot(height=t(as.matrix(data.frame(test.rand))), beside = F, las=2, col=c(mypal[1:5], "white",mypal[6]), ylab="% Variance Explained",
         names.arg = c("TLP","SWA","C[T]","TWL"),)
 legend(xpd=NA, x = 4.5, y=0.7, legend=rev(colnames(test.rand)), fill=rev(c(mypal[1:5], "white",mypal[6])), ncol=1, bty="n",  cex=1)



```
\
Sooo, I get somewhat different results with rarified vs unrarified fullpvs dataset (rarifying down to 34 Sequoiadendron giganteum and Pinus albicaulis). But in general, it's actually not that different, either in what dredge says is important or what the var decomp says.
\
However, the answer is substantially different for some traits with a (1|Species) random effect vs no species random effect. I think I prob need the species, but maybe I can rerun on the fullspp and get the same thing?
\
- But for both TLP and SWA pretty much all the within species var gets thrown into Life.Stage




\
# Putting it together: PV curves we can 'see' from RS
This is the actual relationship between water potential and absolute g water content (per m2 leaf area). 
This assumes identical LAI or the viewing of a single leaf. 

``` {r plotting raw Psi per g curves, echo=F}
mypal <- brewer.pal(n=8, "Dark2")
mypal.light <- paste0(brewer.pal(n=8, "Dark2"),"66")
palette(mypal.light)
plot(TLP~SWA, fullpvs, type="n"
     , xlim=c(min(pvs$LWAtlp,na.rm=T), max(pvs$SWA,na.rm=T))
     , ylim=c(min(pvs$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="g water per m2 leaf")
# for(i in which(pvs$Species=="Sequoiadendron giganteum")){
#   lines(x=c(pvs$SWA[i],pvs$LWAtlp[i]), y=c(0,pvs$TLP[i]), col="green")
# }
for(i in 1:nrow(fullpvs)){
  lines(x=c(fullpvs$SWA[i],fullpvs$LWAtlp[i]), y=c(0,fullpvs$TLP[i]), col= factor(fullpvs$Biome)[i]) #grey= "#11111144")
}
legend("bottom",legend=levels(factor(fullpvs$Biome)), lty=1, col=mypal.light, lwd=3,ncol=3, bty="n")
mtext(text=paste("N=",length(which(fullpvs$water_loss.m2>0)), ";  # Species=",length(unique(fullpvs$Species[which(fullpvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)


```


``` {r psi~ g curve for spp}
quartz(width=5.5, height=4)
mypal <- brewer.pal(n=8, "Dark2")
mypal.light <- paste0(brewer.pal(n=8, "Dark2"),"66")
palette(mypal.light)
plot(TLP~SWA, fullspp, type="n"
     , xlim=c(min(pvs$LWAtlp,na.rm=T), max(fullspp$SWA,na.rm=T))
     , ylim=c(min(pvs$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="g water per m2 leaf", yaxs="i")
# for(i in which(pvs$Species=="Sequoiadendron giganteum")){
#   lines(x=c(pvs$SWA[i],pvs$LWAtlp[i]), y=c(0,pvs$TLP[i]), col="green")
# }
#palette(mypal)
for(i in 1:nrow(fullspp)){
  lines(x=c(fullspp$SWA[i],fullspp$LWAtlp[i]), y=c(0,-1*fullspp$TLP[i]), col= factor(fullspp$Biome)[i], lwd=1.2) #grey= "#11111144")
}
legend("bottomright",legend=levels(factor(fullspp$Biome)), lty=1, col=mypal, lwd=1.5,ncol=1, bty="n", cex=.7)
mtext(text=paste( "# Species=",length(unique(fullpvs$Species[which(fullpvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)
for(i in 1:nrow(fullspp[which(fullspp$Gymno.Angio=="Gymnosperm"),])){
  lines(x=c(fullspp[which(fullspp$Gymno.Angio=="Gymnosperm"),]$SWA[i],fullspp[which(fullspp$Gymno.Angio=="Gymnosperm"),]$LWAtlp[i]), y=c(0,-1*fullspp[which(fullspp$Gymno.Angio=="Gymnosperm"),]$TLP[i]), col="black", lwd=2) #grey= "#11111144")
}
legend("bottomleft", legend = "Gymnosperms", lwd=2, col="black", bty="n", cex=.9)
```
And using giant sequoia as a case study for how much this can vary within a species (since we have 440 values from Sequoiadendron giganteum - Thank you Cam Williams!)

``` {r within species sequoia variation, echo=F}
plot(TLP~SWA, fullpvs, type="n"
     , xlim=c(0, max(fullpvs$SWA,na.rm=T)) # min(fullpvs$LWAtlp,na.rm=T)-100
     , ylim=c(min(fullpvs$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="g water per m2 leaf")
for(i in which(fullpvs$Species=="Sequoiadendron giganteum")){
  lines(x=c(fullpvs$SWA[i],fullpvs$LWAtlp[i]), y=c(0,fullpvs$TLP[i]), col="green")
}
for(i in 1:nrow(fullpvs)){
  lines(x=c(fullpvs$SWA[i],fullpvs$LWAtlp[i]), y=c(0,fullpvs$TLP[i]), col="#11111144")
}
mtext(text=paste("N=",length(which(fullpvs$water_loss.m2>0)), ";  # Species=",length(unique(fullpvs$Species[which(fullpvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)

legend('bottom', legend="Sequoiadendron",col="green",lwd=2, bty="n")

```



Let's look at the 7 species that have at least 10 curves in the dataset as well

``` {r within species variation, echo=F}
quartz(width=5.5, height=4)
mypal <- brewer.pal(n=8, "Dark2")
mypal.light <- paste0(brewer.pal(n=8, "Dark2"),"66")
palette(mypal.light)

plot(TLP~SWA, fullpvs, type="n"
     , xlim=c(min(pvs$LWAtlp,na.rm=T), max(fullspp$SWA,na.rm=T))
     , ylim=c(min(pvs$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="LWA (g water per m2 leaf)", yaxs="i")
rep.spp <- names(xtabs(~Species, fullpvs)[which(xtabs(~Species, fullpvs)>10)][order(xtabs(~Species, fullpvs)[which(xtabs(~Species, fullpvs)>10)], decreasing=T)])
rep.spp.names <- c("Seq. giganteum","Pinus albicaulis","Quercus garryana","Pinus longaeva","Davilla kunthii", "Terminalia amazonia", "Zea mays")
# for(i in 1:nrow(fullpvs)){
#   lines(x=c(fullpvs$SWA[i],fullpvs$LWAtlp[i]), y=c(0,fullpvs$TLP[i]), col="#11111144")
# }
for(i in 1:nrow(fullspp)){
  lines(x=c(fullspp$SWA[i],fullspp$LWAtlp[i]), y=c(0,-1*fullspp$TLP[i]), col= "#11111144", lwd=1) #grey= "#11111144")
}
palette(mypal)
for(spp in 1:length(rep.spp)){
  tmp <- fullpvs %>% filter(Species == rep.spp[spp])
  for(i in 1:nrow(tmp)){
    lines(x=c(tmp$SWA[i],tmp$LWAtlp[i]), y=c(0,tmp$TLP[i]), col=spp)
    }
}

legend(x=-190, y=1.5, legend = rep.spp.names, lwd=2, col=mypal[1:7], xpd=NA, ncol=4, cex=.8, bty="n")

#mtext(text=paste("N=",length(which(fullpvs$water_loss.m2>0)), ";  # Species=",length(unique(fullpvs$Species[which(fullpvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)


```

#### Gymnosperms Only
``` {r gymnos only}
gymnos <- fullpvs[which(fullpvs$Gymno.Angio=="Gymnosperm"),]
plot(TLP~SWA, gymnos, type="n"
     , xlim=c(min(gymnos$LWAtlp,na.rm=T), max(gymnos$SWA,na.rm=T))
     , ylim=c(min(gymnos$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="g water per m2 leaf")
rep.spp <- names(xtabs(~Species, gymnos)[which(xtabs(~Species, gymnos)>10)])
for(i in 1:nrow(gymnos)){
  lines(x=c(gymnos$SWA[i],gymnos$LWAtlp[i]), y=c(0,gymnos$TLP[i]), col="#11111144")
}

for(spp in 1:length(rep.spp)){
  tmp <- gymnos %>% filter(Species == rep.spp[spp])
  for(i in 1:nrow(tmp)){
    lines(x=c(tmp$SWA[i],tmp$LWAtlp[i]), y=c(0,tmp$TLP[i]), col=spp, lwd=1)
    }
}

mtext(text=paste("N=",length(which(gymnos$water_loss.m2>0)), ";  # Species=",length(unique(gymnos$Species[which(gymnos$water_loss.m2>0)]))), side=3,line = 0, adj = 0)

```

#### Angiosperms Only (note change in x axis)
``` {r angios only}
angios <- fullpvs[which(fullpvs$Gymno.Angio=="Angiosperm" & fullpvs$SWA<800),]
plot(TLP~SWA, angios, type="n"
     , xlim=c(min(angios$LWAtlp,na.rm=T), max(angios$SWA,na.rm=T))
     , ylim=c(min(angios$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="g water per m2 leaf")
rep.spp <- names(xtabs(~Species, angios)[which(xtabs(~Species, angios)>10)])
for(i in 1:nrow(angios)){
  lines(x=c(angios$SWA[i],angios$LWAtlp[i]), y=c(0,angios$TLP[i]), col="#11111144")
}

for(spp in 1:length(rep.spp)){
  tmp <- angios %>% filter(Species == rep.spp[spp])
  for(i in 1:nrow(tmp)){
    lines(x=c(tmp$SWA[i],tmp$LWAtlp[i]), y=c(0,tmp$TLP[i]), col=spp, lwd=1.4)
    }
}

mtext(text=paste("N=",length(which(angios$water_loss.m2>0)), ";  # Species=",length(unique(angios$Species[which(angios$water_loss.m2>0)]))), side=3,line = 0, adj = 0)

```

So without digging in too much, this confirms the results of the variance decompositions:\
- within species variation appears LARGE in conifers (as is variation in saturated water content per m2)\
- within species variation is perhaps smaller (though not insubstantial) in angiosperms, but angios just have less variation in SWC m-2 than those pesky gymnos

\
\
## Getting rid of effect of SWA
```{r just CFT}

fullpal <- palette()
fullpallight <- paste0(fullpal[1:7], "66")
fullpallight[2] <- "#D95F02"
palette(fullpallight)
plot(TLP*-1~I(-1*water_loss.m2), fullspp, col=factor(Gymno.Angio), xlim=c(-1*max(fullspp$water_loss.m2, na.rm=T),0), pch=16, cex=.6, ylim=c(-1*max(fullspp$TLP,na.rm=T),0), xlab="g water lost from saturation (per m2 leaf)")
mtext(side=3, "PV curve as g water lost")
for(i in 1:nrow(fullspp)){
  lines(x=c(0,-1*fullspp$water_loss.m2[i]), y=c(0,-1*fullspp$TLP[i]), col=factor(fullspp$Gymno.Angio)[i])
}
```
# Figures for AGU


\
\
# __________________________________________________________
# And the Methodological elephant in the room: Rehydration


Using data from Hahm et al. 2018 Ecosphere, where they measured PV curves on Quercus garyana using leaves that were rehydrated versus non rehydrated.

#### Reconstructing PV curves (i.e. in RWC space)
``` {r rehydration data and plot, echo=F}
# Data from Hahm 2018 Ecosphere
qgar <- read.csv("data/params/Hahm_2018_Ecosphere_Quercusgarryana_20210115.csv")
  # a number of empty rows came in at the end, so kill those
qgar <- qgar[which(!is.na(qgar$Tree.ID)),]
  # relevel the 'rehydration' column so it doesn't have an empty NA level
qgar$Rehydration <- factor(qgar$Rehydration)
qgar$LMA <- with(qgar, Dry_mass/Fresh_area)

# calculate the g of h20 per m2 of leaf area at full saturation
qgar$SWA <- qgar$SWC * qgar$LMA
# calculate the g of h20 per m2 leaf area at tlp
qgar$LWAtlp <-qgar$SWA * qgar$RWC_tlp
# calculate the water lost per m2 leaf area between saturation and tlp
qgar$water_loss.m2 <- with(qgar, SWA - LWAtlp)


mypal2 <- brewer.pal(n=8, "Set1")
mypal2.light <- paste0(mypal2, "66")
palette(mypal2.light)

### Use the Quercus garyana dataset to look at how much rehydrating leaves before making PV curve measurements matters
## RWC
plot(TLP~RWC_tlp, qgar, col=Rehydration, ylim=c(-5,0),xlim=c(.50,1.00), pch=16, ylab="Water Potential (MPa)", xlab="%RWC")
for(i in 1:nrow(qgar)){
  lines(x=c(1,qgar$RWC_tlp[i]), y=c(0,qgar$TLP[i]), col=qgar$Rehydration[i])
}
legend('topleft', legend=c("unrehydrated", 'rehydrated'), col=mypal2.light[c(1,2)], pch=16)
mtext("Method:Rehyd vs no rehyd")
```

### Now in absolute water (g per m-2 leaf area) space
``` {r absolute water}
##Absolute g water
plot(TLP~ I(-1*water_loss.m2), qgar, col=Rehydration, ylim=c(-5,0),xlim=c(-140,0), pch=16, ylab="Water Potential (MPa)", xlab="g water loss from saturation (per m2 leaf)")
for(i in 1:nrow(qgar)){
  lines(x=c(0,-1*qgar$water_loss.m2[i]), y=c(0,qgar$TLP[i]), col=qgar$Rehydration[i])
}
legend('topleft', legend=c("unrehydrated", 'rehydrated'), col=mypal2[c(1,2)], pch=16)
mtext("Method:Rehyd vs no rehyd")


```
\
still a significant problem with getting different answers if you rehydrate leaves versus measure from field collection water potential...\
-sure, it doesn't necessarily change the TLP enormously. But BOY does it change the dehydration dynamics...

```{r testing influences, include=FALSE}
tmptwl <- fullpvs[complete.cases(fullpvs %>% select(water_loss.m2,Gymno.Angio,Family,Genus,Species,Wild.Garden,Life.Stage,Biome)),]
modTWL <- lm(log(water_loss.m2)~Gymno.Angio + Family + Genus + Species + Wild.Garden + Life.Stage + Biome,  tmptwl, na.action = na.fail)

modTWL <- lm(log(water_loss.m2)~Gymno.Angio + Species + Wild.Garden + Life.Stage + Biome,  tmptwl, na.action = na.fail)


bestTWL <- dredge(modTWL)

```



\
\
\

# __________________________________________________________
# Stream of Conciousness : Other plots less thought through...

### SWC,LMA and TLP: test of some previously used correlations ###
```{r more plots, include=T}
par(mfrow=c(2,2), mar=c(3.5,0,1,0), oma=c(0,4,0,1), mgp=c(2.5,1,0))
plot(SWC~TLP, fullspp, col=factor(Biome), pch=16, log="xy", xlab="TLP (MPa)", cex.lab=1.1, ylab=expression(paste(SWC[mass])))
abline(lm(log(SWC,10)~I(log(TLP,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWC,10)~I(log(TLP,10)), fullspp))$r.squared,2)), side = 1, line=-1.5, adj = .2)
mtext(expression(paste(SWC[mass])), side=2, line=2.5)

plot(SWC~LMA, fullspp, col=factor(Biome), pch=16, log="xy", xlab="LMA", cex.lab=1.1, yaxt="n")
abline(lm(log(SWC,10)~I(log(LMA,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWC,10)~I(log(LMA,10)), fullspp))$r.squared,2)), side = 1, line=-1.5, adj = .2)

plot(LMA~TLP, fullspp, col=factor(Biome), pch=16, log="xy", xlab="TLP (MPa)", cex.lab=1.1, ylab="LMA")
abline(lm(log(LMA,10)~I(log(TLP,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(LMA,10)~I(log(TLP,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
mtext("LMA", side=2, line=2.5)

```

Can we predict a leaf's characteristics from other characteristics of it? (when those characteristics aren't mathmatically related)
- Not really. 
- The previously hypothesized correlation between TLP and SWC/LDMC isn't great




#### Exapnded biplots: Moving towards a fishing expedition ###

```{r more plotz}
par(mfrow=c(2,5), mar=c(3.5,0,1,0), oma=c(0,4,0,1), mgp=c(2.5,1,0))
plot(water_loss.m2~CFT.g.m2, fullspp, col=factor(Biome), pch=16, log="xy", xlab=expression(paste("Capacitance ", g*MPa*m^2)))
abline(lm(log(water_loss.m2,10)~I(log(CFT.g.m2,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(water_loss.m2,10)~I(log(CFT.g.m2,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
mtext(expression(paste(Delta,"Water ",m^-2)), side=2, line=2.5)
plot(water_loss.m2~totRWCtlp, fullspp, col=factor(Biome), pch=16, log="xy", xlab="%RWC @ TLP", yaxt="n")
abline(lm(log(water_loss.m2,10)~I(log(totRWCtlp,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(water_loss.m2,10)~I(log(totRWCtlp,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
plot(water_loss.m2~SWC, fullspp, col=factor(Biome), pch=16, log="xy", xlab="SWC gh20 gdrymass", yaxt="n")
abline(lm(log(water_loss.m2,10)~I(log(SWC,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(water_loss.m2,10)~I(log(SWC,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
plot(water_loss.m2~LMA, fullspp, col=factor(Biome), pch=16, log="xy", xlab="LMA g per m2", yaxt="n")
abline(lm(log(water_loss.m2,10)~I(log(LMA,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(water_loss.m2,10)~I(log(LMA,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
plot(water_loss.m2~TLP, fullspp, col=factor(Biome), pch=16, log="xy", xlab="TLP (MPa)", yaxt="n")
abline(lm(log(water_loss.m2,10)~I(log(TLP,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(water_loss.m2,10)~I(log(TLP,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)


plot(SWA~CFT.g.m2, fullspp, col=factor(Biome), pch=16, log="xy", xlab=expression(paste("Capacitance ", g*MPa*m^2)))
abline(lm(log(SWA,10)~I(log(CFT.g.m2,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(CFT.g.m2,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
mtext(expression(paste("Sat gWater ",m^-2,"")), side=2, line=2.5)
plot(SWA~totRWCtlp, fullspp, col=factor(Biome), pch=16, log="xy", xlab="%RWC @ TLP", yaxt="n")
abline(lm(log(SWA,10)~I(log(totRWCtlp,10)), fullspp), lty=2)
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(totRWCtlp,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
plot(SWA~SWC, fullspp, col=factor(Biome), pch=16, log="xy", xlab="SWC gh20 gdrymass", yaxt="n")
abline(lm(log(SWA,10)~I(log(SWC,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(SWC,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
plot(SWA~LMA, fullspp, col=factor(Biome), pch=16, log="xy", xlab="LMA g per m2", yaxt="n")
abline(lm(log(SWA,10)~I(log(LMA,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(LMA,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
plot(SWA~TLP, fullspp, col=factor(Biome), pch=16, log="xy", xlab="TLP (MPa)", yaxt="n")
abline(lm(log(SWA,10)~I(log(TLP,10)), fullspp), lty=2)
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(TLP,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)

```

For some reason Picea mariana is way high for both of these, 
- I can't tell why. It's not extreme. 
It's got a pretty high (2nd highest) saturated water content per m2 but also a high LWAtlp
pretty high LMA, particularly for its SWC?


#####* A few Quick takeaways: ######
1) total water to lose is correlated with the capacitance
2) but not correlated with TLP (the end point)
3) total water to loss is only vaguely correlated with SWC
4) total water to loss is somewhat (but not crazy strongly) correlated with LMA
5) total water to loss is strongly nevatively correlated with totRWCtlp



```{r plots, echo=F}

par(mfrow=c(2,2), mar=c(3.5,0,1,0), oma=c(0,4,0,1), mgp=c(2.5,1,0))
plot(SWC~CFT.g.m2, fullspp, col=factor(Biome), pch=16, log="xy", xlab=expression(paste("Capacitance (",MPa^-1,m^-2,")")), cex.lab=1.1)
abline(lm(log(SWC,10)~I(log(CFT.g.m2,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWC,10)~I(log(CFT.g.m2,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
mtext(expression(paste("Sat Water ",Content[mass])), side=2, line=2.5)

plot(SWC~water_loss.m2, fullspp, col=factor(Biome), pch=16, log="xy", xlab=expression(paste("Total Water Loss (",m^-2,")")), cex.lab=1.1, yaxt="n")
abline(lm(log(SWC,10)~I(log(water_loss.m2,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWC,10)~I(log(water_loss.m2,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)

plot(SWA~CFT.g.m2, fullspp, col=factor(Biome), pch=16, log="xy", xlab=expression(paste("Capacitance (",MPa^-1,m^-2,")")), cex.lab=1.1)
abline(lm(log(SWA,10)~I(log(CFT.g.m2,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(CFT.g.m2,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)
mtext(expression(paste("Sat Water ",Content[area])), side=2, line=2.5)

plot(SWA~water_loss.m2, fullspp, col=factor(Biome), pch=16, log="xy", xlab=expression(paste("Total Water Loss (",m^-2,")")), cex.lab=1.1, yaxt="n")
abline(lm(log(SWA,10)~I(log(water_loss.m2,10)), fullspp))
mtext(paste("R2=",round(summary(lm(log(SWA,10)~I(log(water_loss.m2,10)), fullspp))$r.squared,2)), side = 3, line=-1.5, adj = .2)


```


Total water content (mass basis) not strongly linked to Capacitance or Tot Water Loss (area basis)
But Total Water Content (area basis) is more related to TWL than 



### How much water can they lose?

``` {r how much water can they lose, echo=F}
par(mar=c(8,5,2,1))
boxplot(water_loss.m2~Biome, pvs, las=2, ylim=c(0,200), ylab="g water lost\nfull sat to tlp, per m2 leaf", xlab="")
  #note: there's one value of ~480, which should be investigated
mtext(text=paste("N=",length(which(pvs$water_loss.m2>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)

boxplot(CFT.g.m2~Biome, pvs, las=2, ylab=expression(paste("g water lost ", MPa^-1, m^-2)), xlab="")
mtext(text=paste("N=",length(which(pvs$CFT.g.m2>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$CFT.g.m2>0)]))), side=3,line = 0, adj = 0)

```




TAKEHOMES: 

- conifers have a lot more water to lose per m2 leaf because they've got very high LMA and maybe fairly negative TLPs (difference between water loss per MPa and total water loss from saturation to TLP is the TLP itself).

- Crops are super wet, but generally wet and high capacitance, but have wimpy TLPs so don't have that much water to lose

- Lots going on in the tropics (compare wet and dry and capacitance to SWC above)



``` {r code snippets for mb mid processing, include=F}
# for(i in 1:nrow(mb12clean)){
#      lines(x=c(mb12clean$SWA[i],mb12clean$LWAtlp[i]), y=c(0,mb12clean$TLP[i]), col="red")
#  }
# 
# treats <- factor(mb14wc$Treatment)
# for(i in 1:nrow(mb14wc)){
#      lines(x=c(mb14wc$SWA[i],mb14wc$LWAtlp[i]), y=c(0,mb14wc$TLP[i]), col="blue")
# }
# 
# for(i in 1:nrow(mb14wc)){
#      lines(x=c(mb14wc$SWA[i],mb14wc$LWAtlp[i]), y=c(0,mb14wc$TLP[i]), col="purple")
# }

```


### *Pulling out effect of SWC to just look at slopes*

``` {r drydown in RWC vs Absolute WC space, echo=F}

# RWC plot for Bartlett 2012

plot(TLP~RWC_tlp, mb12, col=Biome, xlim=c(50,100), ylim=c(min(pvs$TLP,na.rm=T),0), pch=16, cex=.8, xlab="%RWC")
mtext(side=3, "PV curves as % of saturated content (Bartlett 2012 dataset)")
for(i in 1:nrow(mb12)){
  lines(x=c(100,mb12$RWC_tlp[i]), y=c(0,mb12$TLP[i]), col=mb12$Biome[i])
}
legend('topleft', cex=.8, bty='n', legend = c("TropicalDry","Temperate"), pch=16, col=mypal.light[1:2])



plot(TLP~totRWCtlp, pvs, col=factor(Biome), xlim=c(50,100), ylim=c(min(pvs$TLP,na.rm=T),0), pch=16, cex=.8, xlab="%RWC")
mtext(side=3, "PV curve as % of saturated content (current dataset)")
for(i in 1:nrow(pvs)){
  lines(x=c(100,pvs$totRWCtlp[i]), y=c(0,pvs$TLP[i]), col=factor(pvs$Biome)[i])
}
#legend('topleft', cex=.8, bty='n', legend = c("TropicalDry","Temperate"), pch=16, col=mypal1light[1:2])




plot(TLP~I(-1*water_loss.m2), pvs, col=factor(Biome), xlim=c(-500,0), ylim=c(min(pvs$TLP,na.rm=T),0), pch=16, cex=.8, xlab="g water lost from saturation (per m2 leaf)")
mtext(side=3, "PV curve as g water lost")
for(i in 1:nrow(pvs)){
  lines(x=c(0,-1*pvs$water_loss.m2[i]), y=c(0,pvs$TLP[i]), col=factor(pvs$Biome)[i])
}

```

Actually a lot more interpretable in terms of g water rather than % RWC!


### OLD: ALL IN RWC SPACE
Exploratory visuals of four types of questions:

1) do biomes/land cover types differ?
2) how much does within-species seasonal adjustment matter?
3) how much does method (measuring rehydrated leaves vs measuring unrehydrated leaves) matter?
4) How appropriate is a linear fit for actual data?

``` {r quick plots RWC~TLP}

mypal1 <- brewer.pal(n=8, "Set1")
mypal1light <- paste0(mypal1, "66")
palette(mypal1light)


plot(TLP~RWC_tlp, mb12, col=Biome, xlim=c(68,100), ylim=c(min(mb12$TLP,na.rm=T),0), pch=16, cex=.8, xlab="%RWC")
mtext(side=3, "Global Meta-analysis")
for(i in 1:nrow(mb12)){
  lines(x=c(100,mb12$RWC_tlp[i]), y=c(0,mb12$TLP[i]), col=mb12$Biome[i])
}
legend('topleft', cex=.8, bty='n', legend = c("TropicalDry","Temperate"), pch=16, col=mypal1light[1:2])

plot(TLPdry~RWC_tlpdry, mb14w, col=mypal[1], xlim=c(60,100), ylim=c(min(mb14w$TLPdry,na.rm=T),0), pch=16, cex=.8, ylab="TLP", xlab="%RWC")
for(i in 1:nrow(mb14w)){
  lines(x=c(100,mb14w$RWC_tlpdry[i]), y=c(0,mb14w$TLPdry[i]), col=paste0(mypal[1],"33"))
}

points(TLPwet~RWC_tlpwet, mb14w, col=mypal[2], pch=16, cex=.8)
for(i in 1:nrow(mb14w)){
  lines(x=c(100,mb14w$RWC_tlpwet[i]), y=c(0,mb14w$TLPwet[i]), col=paste0(mypal[2],"33"))
}
arrows(x0=mb14w$RWC_tlpwet, x1=mb14w$RWC_tlpdry, y0=mb14w$TLPwet, y1=mb14w$TLPdry, length = .05)
legend('topleft', cex=.8, bty='n', legend = c("Dry Season","Wet Season"), pch=16, col=mypal1[1:2])
mtext("Seasonal Adjustment")



### Appropriateness of linear fit when you have actual data
plot(WP~RWC, hf, ylab="Water Potential (MPa)", xlab="Relative Water Content", col=type)
legend('topleft', legend = c("Leaf","Leaf+Branch"), col=c(1,2), pch=1)
abline(h=-1.68, lty=2)
text(y=-2.1, x=3, "TLP\n(-1.68 MPa)")
for(i in unique(hf$sample)){
  lines(WP~RWC, hf[which(hf$sample==i),], col=type)
}
abline(lm(WP~RWC, hf[which(hf$WP> -1.68),]), col="black")
mtext("Linear Fit?")

```