---
title: "PV-for-RS_Analysis"
author: "LDL Anderegg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(RColorBrewer)
library(dplyr)
library(reshape2)
mypal <- brewer.pal(n=8, "Set1")
palette(mypal)
require(ggplot2)
require(maps)
library(tidyverse)
library(car)
library(lme4)
library(lmerTest)
library(MuMIn)
```

## The Contributed Dataset:

``` {r import contributed data, echo=F}

  # download protocol: Download from Google Drive, name xlsx version. Remove rows 1,2, and 4 and save as versioned .csv.

# first round of data submissions
pvs1 <- read.csv("data/ContributedData/PV_for_RS_metaanalysis_contributedby2023.12.15.csv", header=T, na.strings = "", )
# clean out extra rows and extra columns
pvs1 <- pvs1[which(!is.na(pvs1$Contributor)),-grep("^X", colnames(pvs1))]
#pvs <- read.csv("data/ContributedData/PV_for_RS_metaanalysis_20231121.csv", header=T, na.strings = "", )
#pvs <- read.csv("data/ContributedData/PV_for_RS_metaanalysis_20220716.csv", header=T, na.strings = "", )

# second round of data submissions
pvs2 <- read.csv("data/ContributedData/PV_for_RS_metaanalysis_2024datacontributions_202400807.csv", header=T, na.strings = "", )[-c(2:5),] # get rid of first 4 demo rows
# clean out extra rows
pvs2 <- pvs2[which(!is.na(pvs2$Contributor)),-grep("^X", colnames(pvs2))]



# rename the column names to get rid of weird characters and stuff
colnames(pvs1) <- c("Contributor","Species","Treatment","Co.auth"
                   ,"Biome","Growth.Form","T.S.L.H","Ever.Decid"
                   ,"Wild.Garden","Life.Stage"
                   ,"TLP","TLS.sd","TLP.se"
                   , "totRWCtlp", "totRWCtlp.sd","totRWCtlp.se"
                   , "LMA", "LMA.sd","LMA.se"
                   ,"SWC", "SWC.sd", "SWC.se"
                   ,"CFT.mol.m2","CFT.sd","CFT.se","Po","E","ApoFrac"
                   ,"sympRWCtlp","Psi.PD","Psi.min","Lat","Lon", "Elev"
                   ,"Site.Name","Rehydrated","Month","Year","OtherEnv"
                   ,"Methods.Notes","Ref","Other.Notes")

colnames(pvs2) <- c("Contributor","Species","Treatment","Co.auth"
                   ,"Biome","Growth.Form","T.S.L.H","Ever.Decid"
                   ,"Wild.Garden","Life.Stage"
                   ,"TLP","TLS.sd","TLP.se"
                   , "totRWCtlp", "totRWCtlp.sd","totRWCtlp.se"
                   , "LMA", "LMA.sd","LMA.se"
                   ,"SWC", "SWC.sd", "SWC.se"
                   ,"CFT.mol.m2","CFT.sd","CFT.se","Po","E","ApoFrac"
                   ,"sympRWCtlp","Psi.PD","Psi.min","Lat","Lon", "Elev"
                   ,"Site.Name","Rehydrated","Month","Year","OtherEnv"
                   ,"Methods.Notes","Ref","Other.Notes")

# combine submission batches
pvs <- rbind(pvs1, pvs2)
pvs$LMA <- as.numeric(pvs$LMA)
pvs$Lat <- as.numeric(pvs$Lat)
pvs$Lon <- as.numeric(pvs$Lon)



# fix LMA values that are in g cm2
pvs$LMA[which(pvs$LMA<1)] <- pvs$LMA[which(pvs$LMA<1)] * 100 * 100
# fix expanded Biomes
pvs$Biome[which(pvs$Biome=="TemperateConifer (but could be montane conifer…)")] <- "TemperateConifer"
pvs$Biome[which(pvs$Biome=="semiarid grassland (MAP 397 mm, MAT 8 C)")] <- "Grassland"
pvs$Biome[which(pvs$Biome=="Tropical montane cloud forest")] <- "TropicalMoist"
pvs$Biome[which(pvs$Biome=="Crop ")] <- "Crop"
pvs$Biome[which(pvs$Biome=="Coastal")] <- "Med/TempDry"


pvs$Life.Stage[which(pvs$Life.Stage=="Se (but see Age column)")] <- "Se" # making 3yr olds into seedlings (<1.3m tall)
pvs$Ever.Decid[which(pvs$Ever.Decid=="Dd/Semi-Deciduous")] <- "DD" # currenlty calling semi-deciduous full drought deciduous

# fix T.S.L.H before I added herbaceous
pvs$T.S.L.H[which(pvs$Growth.Form=="H")] <- "H"


# fix TLPs that are positive rather than negative
pvs$TLP[which(pvs$TLP>0)] <- -1 * pvs$TLP[which(pvs$TLP>0)]

# Some totRWC_tlp are VERY weird. 
# in MB12 there aren't any <70% (but only 89)
# in mb14w there 3 (dry) and 4 (wet) RWC_tlps lower than 70, but only 1 lower than 50 (wet)
# and nothing <60% for mb14c
#hist(pvs$totRWCtlp,breaks=50)
#hist(mb14w$RWC_tlpwet, col="red", add=T)
#hist(mb12$RWC_tlp, col="blue", add=T)
#hist(mb14c$RWC_total_wet_percent, col="green", add=T)

# everything below 50 seems to be either a typo or bad.
pvs$Other.Notes[which(pvs$totRWCtlp<50)] <- paste("BAD totRWCtlp", pvs$Other.Notes[which(pvs$totRWCtlp<50)] )
#pvs$totRWCtlp.se[which(pvs$totRWCtlp<50)] # and one has a massive SD so prob bad curves?
pvs$totRWCtlp[which(pvs$totRWCtlp<50)] <- NA 
# LMAs look good (lots of high ones, but those are Sequoiadendron, which scans with TRY)
# SWC - two look unusually high (>10). but according to TRY those aren't SUPER crazy, just pretty crazy. One is the Eugenia procera with -13% totRWCtlp though, so maybe just remove this one?
  # on further exploration, they're both probably bad. They're normalish plants and all TRY extreme SWC plants are majore weirdos.
pvs$Other.Notes[which(pvs$SWC>10)] <- paste("BAD SWC,", pvs$Other.Notes[which(pvs$SWC>10)])
pvs$SWC[which(pvs$SWC>10)] <- NA
# switch osmotic potential to be negative
pvs$Po[which(pvs$Po>0)] <- -1*pvs$Po[which(pvs$Po>0)]
# palms/ferns have high E in mb12 (60-80), so values 60-90 don't look like obvious bummers.
# but E>300 for Forsteronia spicata seems like a data error
pvs$Other.Notes[which(pvs$E>300)] <- paste("BAD ModE,", pvs$Other.Notes[which(pvs$E>300)])
pvs$E[which(pvs$E>300)] <- NA
# Apoplastic Fraction: mb12 median ~0.2
  # but we've got some negative ones and some >1s. even some >100s. I think I'll have to treat all of them as bad and double check with Danielle Ulrich about them
  # negative numbers are all from oaks and noted as bad
pvs$Other.Notes[which(pvs$ApoFrac>1 | pvs$ApoFrac<0)] <- paste("BAD ApoFrac,",pvs$Other.Notes[which(pvs$ApoFrac>1 | pvs$ApoFrac<0)] )
pvs$ApoFrac[which(pvs$ApoFrac>1 | pvs$ApoFrac<0)] <- NA
# clean rehydrated method flag
pvs$Rehydrated[which(pvs$Rehydrated=="no")] <- "N"
pvs$Rehydrated[which(pvs$Rehydrated=="yes")] <- "Y"
pvs$Rehydrated[which(pvs$Rehydrated=="Yes")] <- "Y"
#pvs$Rehydrated[which(pvs$Rehydrated=="no (predawn)")] <- "PD


```

*Data quality notes: To Double Check
Danielle Ulrich - apoplastic fractions >1, German Vargas - Species with crazy totRWCtlp and other values,Chris M. Smith-Martin - CFT calculation off.




``` {r calculate some desired quantities}
# calculate the g of h20 per m2 of leaf area at full saturation
pvs$SWA <- pvs$SWC * pvs$LMA
# calculate the g of h20 per m2 leaf area at tlp
pvs$LWAtlp <-pvs$SWA * pvs$totRWCtlp /100
# calculate the water lost per m2 leaf area between saturation and tlp
pvs$water_loss.m2 <- with(pvs, SWA - LWAtlp)
# turn capacitance into g instead of mol
pvs$CFT.g.m2 <- pvs$CFT.mol.m2 * 18.01528
  #NOTE: Some of these are VERY different than calcualting it myself... (water_loss.m2/(-1*TLP))
pvs$CFT.g.m2_mycalc <- pvs$water_loss.m2/(-1*pvs$TLP)
#plot(CFT.g.m2_mycalc~CFT.g.m2, pvs, xlim=c(0,2.5), ylim=c(0,20))
  # looks like all points except one with reported CFT<3 are in different units, and 3 points with CFT 4-6 are in different units. but its not a simple mol->g conversion. nor cm2->m2.
  # all of the ones with a unit offset are Chris M. Smith-Martin's will roll with my calc for those.
pvs$CFT.g.m2[which(pvs$Contributor=="Chris M. Smith-Martin")] <- pvs$CFT.g.m2_mycalc[which(pvs$Contributor=="Chris M. Smith-Martin")]
pvs$Other.Notes[which(pvs$Contributor=="Chris M. Smith-Martin")] <- paste("CFT_recalced,", pvs$Other.Notes[which(pvs$Contributor=="Chris M. Smith-Martin")])

# we could just use my calc for everything except where I can't calc it myself? --> there are 19 CFTs that I have calcs for that aren't contributed and ~80 that are contributed that I can't calc.
pvs$CFT.g.m2[which(pvs$CFT.g.m2_mycalc>0)] <- pvs$CFT.g.m2_mycalc[which(pvs$CFT.g.m2_mycalc>0)]


```

``` {r summarize dataset}
print(paste("Number of entries in Contributed Data:", nrow(pvs)))
print(paste("Number of species:", length(unique(pvs$Species))))
hist(xtabs(~Species, pvs), breaks=150, xlim=c(0,20), xlab="entries per species", main="")
print(paste("Number of species with >5 entries:", length(which(xtabs(~Species, pvs)>5))))
print(paste("Number of species with >50 entries:", length(which(xtabs(~Species, pvs)>50))))
print(paste("Number of unique 'sites':",length(unique(pvs$Site.Name))))
print("Where traits were measured:")
xtabs(~Wild.Garden ,pvs)
maps::map(database="world", col="black")
points(Lat~Lon, pvs, pch=16, cex=.5, col="green")
mtext("Contributed Curves",side=3)
```


No real representation outside of the Americas at present.

Number of entries with complete trait values:
```{r more summaries, echo=FALSE}
apply(pvs %>% select(TLP,totRWCtlp,LMA,SWC,SWA,LWAtlp,CFT.mol.m2,water_loss.m2,CFT.g.m2,CFT.g.m2_mycalc)
      , MARGIN = 2
      ,FUN = function(x){length(which(!is.na(x)))})
```
-**totRWCtlp** = relative water content at TLP

-**SWC** = g water per g leaf dry mass at full saturation

-**CFT...** = leaf capacitance per m2 leaf area

-**water_loss.m2** = total g water lost per m2 leaf area going from full sutration to TLP (calculated above)

-**SWA** = g water per m2 leaf at full saturation

-**LWCtlp** = g water per m2 leaf area at TLP (LWA at TLP)




### Bring in Data from Bartlett Meta-analyses ###
We'll bring in Megan's published meta-analyses from Bartlett et al. 2012 and Bartlett et al. 2014.
However, these only have RWCtlp and sometimes LMA, and we need SWC (or LDMC) to transform these curves into absolute gH2O per mass or area.
So I pulled all the LMA and LDMC data from TRY to infill with species average data where I can.
```{r import data, include=F}
# data from Bartlett 2012 ELE
mb12raw <- read.csv("data/params/Bartlett2012ELE.csv", header=T) 
  # make clean version with all rows with both TLP and RWC_tlp values
mb12 <- mb12raw[which(mb12raw$TLP<0 & mb12raw$RWC_tlp>0),]
  # relevel things so that empty levels don't use all my colors
mb12$Biome <- factor(mb12$Biome)
mb12$Biome <- relevel(mb12$Biome, ref="Temperate")
mb12$Biome <- relevel(mb12$Biome, ref="TropicalDry")

# Bartlett 2014 ELE wild plants
mb14wraw <- read.csv("data/params/Bartlett2014ELE_wild.csv") # data from 
mb14wraw <- mb14wraw %>% rename(TLPwet=TLP_WET_MPa, TLPdry=TLP_DRY_MPa, RWC_tlpwet=RWC_total_wet_percent, RWC_tlpdry = RWC_total_dry_percent)
mb14w <- mb14wraw[which(mb14wraw$RWC_tlpwet>0),]
mb14c <- read.csv("data/params/Bartlett2014ELE_S3crop.csv")

# Data from Hahm 2018 Ecosphere
qgar <- read.csv("data/params/Hahm_2018_Ecosphere_Quercusgarryana_20210115.csv")
  # a number of empty rows came in at the end, so kill those
qgar <- qgar[which(!is.na(qgar$Tree.ID)),]
  # relevel the 'rehydration' column so it doesn't have an empty NA level
qgar$Rehydration <- factor(qgar$Rehydration)
qgar$LMA <- with(qgar, Dry_mass/Fresh_area)

## Data from Holtzman et al. 2021 from Harvard Forest
hf <- read.csv("data/raw_RWC-Psi/Holtzman_HarvardForest_rawPVcurves_20200304.csv"
)
hf$type <- factor(hf$type)

## example curves from Cam for illustration
ecs <- readxl::read_xlsx(path = "data/ContributedData/ExampleCurves_MeganAndCam.xlsx", sheet = 2, skip = 1)
ecs$g_h20_m2 <- ecs$Mass_h2o/ecs$Fresh_area_m2 # calculate the total water per m2 leaf area (g water / leaf area in m2)

# just verifying that the above works with the actual math
# test <- ecs$SWC * ecs$`Dry mass`/ecs$Fresh_area_m2 * ecs$RWC # it does! good.



```




```{r adding SLA, include=F}
#read in TRY species averages (from Cd-TRY_LMALDMC_processing)
SLAspp <- read.csv("DerivedData/TRY_SpeciesAverages_20230719.csv", header=T, row.names =1 )


# add a SLA column from TRY
mb12$SLA <- SLAspp$SLA[match(mb12$Species, SLAspp$AccSpeciesName)] # fills in 50 spp
# add a column to keep track of which traits got imputed from TRY
mb12$LMA_imputed <- 0
mb12$LMA_imputed[which(is.na(mb12$LMA))] <- 1
mb12$LMA[which(is.na(mb12$LMA))] <- 1/mb12$SLA[which(is.na(mb12$LMA))]*1000  # wohoo! fills in a whopping 20 spp...
mb12$SWC <- SLAspp$SWC[match(mb12$Species, SLAspp$AccSpeciesName)] # fills in 0 spp
mb12$SWC <- SLAspp$SWC_from_LDMC[match(mb12$Species, SLAspp$AccSpeciesName)] # only fills in 24 spp. Dang. But at least all 24 of them have LMA values!
mb12$SWC_imputed <- 1

## assign a single LMA and single SWC to both seasons of mb14 dataset

mb14w$LMA <- 1/SLAspp$SLA[match(mb14w$Species, SLAspp$AccSpeciesName)]*1000 # get 53 spp out of it
mb14w$SWC <- SLAspp$SWC_from_LDMC[match(mb14w$Species, SLAspp$AccSpeciesName)] # get an additional 27 species out of it. All of them with both LMA and SWC
# assign Eucalyptus rubida the mean SWC for non-weird Eucs (there's one study with RIDICULOUSLY high LDMCs in TRY (>0.8, meaning 0.16 SWC), which must be methodological)
eucLDMCs <- SLAspp[c(grep("Eucalyptus",SLAspp$AccSpeciesName), grep("Corymbia", SLAspp$AccSpeciesName)),] %>% filter(LDMC<.7) %>% select(LDMC)

mb14w$SWC[which(mb14w$Species=="Eucalyptus rubida")] <-1/mean(eucLDMCs$LDMC)-1
# but actually only 21 because 6 are missing totRWCtlp
mb14w$LMA_imputed <- 1
mb14w$SWC_imputed <- 1

mb14c$LMA <- 1/SLAspp$SLA[match(mb14c$Species, SLAspp$AccSpeciesName)]*1000  # get 37 spp
mb14c$SWC <- SLAspp$SWC_from_LDMC[match(mb14c$Species, SLAspp$AccSpeciesName)]
  # get 29 species out of it with both LMA and SWC
mb14c$LMA_imputed <- 1
mb14c$SWC_imputed <- 1

# clean up the columns to get ready to merge
mb12clean <- mb12 %>% filter(TLP<0 & RWC_tlp>0 & LMA>0 & SWC>0) %>% select(Species, Biome, "Growth.Form"=GrowthForm, "Ever.Decid"=Ever_Decid, TLP, "totRWCtlp"=RWC_tlp, LMA, SWC,"Po"=Pot_osmot,"E"=Mod_elast,"ApoFrac"=Apo_frac,"sympRWCtlp"=RWC_symp_tlp,"Ref"=Reference, LMA_imputed, SWC_imputed)
mb12clean$Treatment <- NA
mb12clean$T.S.L.H <- NA
mb12clean$Wild.Garden <- NA
mb12clean$Life.Stage <- NA
mb12clean$Psi.PD <- NA
mb12clean$Psi.min <- NA
mb12clean$Rehydrated <- NA
mb12clean$Lat <- NA
mb12clean$Lon <- NA
mb12clean$Elev <- NA
# calculate the g of h20 per m2 of leaf area at full saturation
mb12clean$SWA <- mb12clean$SWC * mb12clean$LMA
# calculate the g of h20 per m2 leaf area at tlp
mb12clean$LWAtlp <-mb12clean$SWA * mb12clean$totRWCtlp /100
# calculate the water lost per m2 leaf area between saturation and tlp
mb12clean$water_loss.m2 <- with(mb12clean, SWA - LWAtlp)
# calculate capacitance, g H2o m-2 MPa-1
mb12clean$CFT.g.m2 <- mb12clean$water_loss.m2/(-1*mb12clean$TLP) #mb12clean$CFT.mol.m2 * 18.01528
mb12clean$Other.Notes <- NA
mb12clean$Treatment <- NA
mb12clean$Dataset <- "mb12"

# process and arrange mb14 wild dataset
mb14wclean <- mb14w %>% filter(TLPwet<0 & RWC_tlpwet>0 & LMA>0 & SWC>0)%>% select(Species, "Biome"=Biomes, "Growth.Form"=Woody_Herbaceous, "Ever.Decid"=Evergreen_Deciduous_Herb, "TLP"=TLPwet, "totRWCtlp"=RWC_tlpwet,"Po"=OsmoticPotential_WET_MPa, LMA, SWC,"Ref"=Reference, LMA_imputed, SWC_imputed, TLPdry, "totRWCtlpdry"=RWC_tlpdry, "Podry"=OsmoticPotential_DRY_MPa, "Lat"=Site_Latitude, "Lon" = Site_Longitude)
# hack together a long form with dry values
mb14wdry <- mb14wclean %>% select(-TLP,-totRWCtlp, -Po)
colnames(mb14wdry)<- str_remove(colnames(mb14wdry),"dry")
mb14wwet <- mb14wclean %>% select(-grep("dry", colnames(mb14wclean)))
mb14wdry <- mb14wdry %>% select(colnames(mb14wwet))
mb14wwet$Treatment <- "Wet"
mb14wdry$Treatment <- "Dry"
mb14wc <- rbind(mb14wwet, mb14wdry)

mb14wc$ApoFrac <- NA
mb14wc$sympRWCtlp <- NA
mb14wc$Treatment <- NA
mb14wc$T.S.L.H <- NA
mb14wc$Wild.Garden <- NA
mb14wc$Life.Stage <- NA
mb14wc$Psi.PD <- NA
mb14wc$Psi.min <- NA
mb14wc$Rehydrated <- NA
mb14wc$Elev <- NA
# calculate the g of h20 per m2 of leaf area at full saturation
mb14wc$SWA <- mb14wc$SWC * mb14wc$LMA
# calculate the g of h20 per m2 leaf area at tlp
mb14wc$LWAtlp <-mb14wc$SWA * mb14wc$totRWCtlp /100
# calculate the water lost per m2 leaf area between saturation and tlp
mb14wc$water_loss.m2 <- with(mb14wc, SWA - LWAtlp)
# calculate capacitance, g H2o m-2 MPa-1
mb14wc$CFT.g.m2 <- mb14wc$water_loss.m2/(-1*mb14wc$TLP) #mb14wc$CFT.mol.m2 * 18.01528
mb14wc$Other.Notes <- NA
mb14wc$Dataset <- "mb14"

# process and arrange mb14 crop dataset
mb14cclean <- mb14c %>% filter(TLP_WET_MPa<0 & RWC_total_wet_percent>0 & LMA>0 & SWC>0) %>% select(Species, "Biome"=Biome, "Growth.Form"=Woody_Herbaceous, "Ever.Decid"=Evergreen_Deciduous_Herb, "TLP"=TLP_WET_MPa, "totRWCtlp"=RWC_total_wet_percent,"Po"=OsmoticPotential_WET_MPa, LMA, SWC,"Ref"=Reference, LMA_imputed, SWC_imputed, "TLPdry"=TLP_DRY_MPa, "totRWCtlpdry"=RWC_total_dry_percent, "Podry"=OsmoticPotential_DRY_MPa, "Other.Notes"=Cultivar)
# hack together a long form with dry values
mb14cdry <- mb14cclean %>% select(-TLP,-totRWCtlp, -Po)
colnames(mb14cdry)<- str_remove(colnames(mb14cdry),"dry")
mb14cwet <- mb14cclean %>% select(-grep("dry", colnames(mb14cclean)))
mb14cdry <- mb14cdry %>% select(colnames(mb14cwet))
mb14cwet$Treatment <- "Wet"
mb14cdry$Treatment <- "Dry"
mb14cc <- rbind(mb14cwet, mb14cdry)

mb14cc$ApoFrac <- NA
mb14cc$sympRWCtlp <- NA
mb14cc$Treatment <- NA
mb14cc$T.S.L.H <- NA
mb14cc$Wild.Garden <- NA
mb14cc$Life.Stage <- NA
mb14cc$Psi.PD <- NA
mb14cc$Psi.min <- NA
mb14cc$Rehydrated <- NA
mb14cc$Lat <- NA
mb14cc$Lon <- NA
mb14cc$Elev <- NA
# calculate the g of h20 per m2 of leaf area at full saturation
mb14cc$SWA <- mb14cc$SWC * mb14cc$LMA
# calculate the g of h20 per m2 leaf area at tlp
mb14cc$LWAtlp <-mb14cc$SWA * mb14cc$totRWCtlp /100
# calculate the water lost per m2 leaf area between saturation and tlp
mb14cc$water_loss.m2 <- with(mb14cc, SWA - LWAtlp)
# calculate capacitance, g H2o m-2 MPa-1
mb14cc$CFT.g.m2 <- mb14cc$water_loss.m2/(-1*mb14cc$TLP) #mb14cc$CFT.mol.m2 * 18.01528
mb14cc$Dataset <- "mb14"

# get everything to have the same column order as mb14wc
mb12clean <- mb12clean %>% select(colnames(mb14wc))
mb14cc <- mb14cc %>% select(colnames(mb14wc))
```

So net result of TRY + mb12 and mb14 datasets is 72 additional species and 120 total values (w and dry for mb14 datasets) I'll take it.


Combining all things into  master dataset, we don't get all the metadata we'd like for all the studies, but we increase the # spp and geographic spread
- Let's analyze the relationship between parameters on the full dataset, and then we can explore specific patterns in depth in the contributed dataset with covariates that we don't have for the full dataset
```{r make master dataset, include=F}
pvs$LMA_imputed <- 0
pvs$SWC_imputed <- 0
pvs$Dataset <- "Contributed"
pvs.master <- pvs %>% select(colnames(mb14wc))
pvs.master$LMA_imputed[which(is.na(pvs.master$LMA))] <- 1
pvs.master$LMA[which(is.na(pvs.master$LMA))] <- 1/SLAspp$SLA[match(pvs.master$Species[which(is.na(pvs.master$LMA))], SLAspp$AccSpeciesName)]*1000
pvs.master$SWC_imputed[which(is.na(pvs.master$SWC))] <- 1
pvs.master$SWC[which(is.na(pvs.master$SWC))] <- SLAspp$SWC_from_LDMC[match(pvs.master$Species[which(is.na(pvs.master$SWC))], SLAspp$AccSpeciesName)]
# fill in missing values with imputed traits
pvs.master$SWA <- pvs.master$SWC * pvs.master$LMA 
#length(which(is.na(pvs$SWA)& pvs.master$SWA>0))# fills in 160 values for all three of these
# calculate the g of h20 per m2 leaf area at tlp
pvs.master$LWAtlp <-pvs.master$SWA * pvs.master$totRWCtlp /100
# calculate the water lost per m2 leaf area between saturation and tlp
pvs.master$water_loss.m2 <- with(pvs.master, SWA - LWAtlp)
# calculate the Capacitance (CFT) for 83 rows without it
pvs.master$CFT.g.m2[which(is.na(pvs.master$CFT.g.m2))] <- pvs.master$water_loss.m2[which(is.na(pvs.master$CFT.g.m2))]/(-1*pvs.master$TLP[which(is.na(pvs.master$CFT.g.m2))])


apply(pvs.master %>% select(TLP,totRWCtlp,LMA,SWC,water_loss.m2,CFT.g.m2), MARGIN = 2 ,FUN = function(x){length(which(!is.na(x)))}) - apply(pvs %>% select(TLP,totRWCtlp,LMA,SWC,water_loss.m2,CFT.g.m2) , MARGIN = 2 ,FUN = function(x){length(which(!is.na(x)))})
  #infilled 220 LMAs, 1 SWC, 160 water_loss.m2 and 199 CFT.g.m2



fullpvs <- rbind(pvs.master, mb12clean, mb14wc, mb14cc)

fullpvs$RWCslope <- fullpvs$TLP*(100- fullpvs$totRWCtlp)


```

# MASTER DATASET SUMMARY
``` {r summarize master dataset}
print(paste("Number of entries:", nrow(fullpvs)))
print(paste("Number of species:", length(unique(fullpvs$Species)))) # added 46 spp
hist(xtabs(~Species, fullpvs), breaks=150, xlim=c(0,20), xlab="entries per species", main="") 
print(paste("Number of species with >5 entries:", length(which(xtabs(~Species, fullpvs)>5)))) # added 4 spp with multiple entries
print(paste("Number of species with >50 entries:", length(which(xtabs(~Species, fullpvs)>50))))
print(paste("Number of unique 'sites':",length(unique(fullpvs$Site.Name))))
print("Where traits were measured:")
xtabs(~Wild.Garden ,fullpvs)
maps::map(database="world", col="black")
locs <- fullpvs %>% group_by(Lat, Lon) %>% summarise(nCurves = n(), nSpecies=length(unique(Species)))
points(Lat~Lon, locs, pch=16, cex=sqrt(nSpecies/pi)/2, col="green")
mtext("Full Dataset (size = n species)")
mtext(paste("plus", length(which(is.na(fullpvs$Lat)))," obs and",length(unique(fullpvs$Species[which(is.na(fullpvs$Lat))])), "spp without locations"), side=1, cex=.8)
```
\
So we've got a ton of coverage in the Americas, and remarkably good spp coverage in the new world tropics. But outside of that we're pretty sparse.



``` {r average to spp, include=F}

fullpvs$Ever.Decid[which(fullpvs$Growth.Form=="H" & is.na(fullpvs$Ever.Decid))] <- "H" # call anything that's not woody H for ever/decid because it's hard to tell for perrenials and grasses and stuff. There are some E's and some "H"s (inherited from mb12).
fullpvs$Ever.Decid[which(fullpvs$Species=="Celtis australis")] <- "CD" # just D in mb12
fullpvs$Ever.Decid[which(fullpvs$Species=="Cordia collococca")] <- "DD" # just D in mb12
fullpvs$Biome[which(fullpvs$Species=="Cordia collococca")] <- "TropicalMoist" # both Dry and Moist
fullpvs$Biome[which(fullpvs$Species=="Eucalyptus rubida")] <- "Med/TempDry" # Temperate in mb12 but Med/TempDry in mb14
fullpvs$Ever.Decid[which(fullpvs$Species=="Paullinia pinnata")] <- "DD" # called E in the garden but DD in the wild... going with DD
#fullpvs[which(fullpvs$Species=="Phaseolus vulgaris"),]
fullpvs$Ever.Decid[which(fullpvs$Species=="Quercus rubra")] <- "CD"
fullpvs$Ever.Decid[which(fullpvs$Species=="Tetracera volubilis")] <- "DD" # called both E and DD in contributed data
fullpvs$Ever.Decid[which(fullpvs$Species=="Trigonia rugosa")] <- "DD" # called both E and DD in contributed data
#fullpvs[which(fullpvs$Species=="Triticum aestivum"),]
fullpvs$Biome[which(fullpvs$Species=="Vitis vinifera")] <- "Crop" # called "Crop" and "Med/TempDry" AND "CD"/"E"/"D"
fullpvs$Ever.Decid[which(fullpvs$Species=="Vitis vinifera")] <- "CD"
# Currently rolling with "CD" for all mb "D" species
fullpvs$Ever.Decid[which(fullpvs$Ever.Decid=="D")] <- "CD" # most of them DO seem CD



# I realized it's probably worth looking at things on a mass vs area basis, so calculating water loss per g leaf as well as m2
fullpvs$water_loss.g <- fullpvs$SWC - fullpvs$totRWCtlp*fullpvs$SWC
  # SWC = gwater/g dry mass @ sat. 

fullspp <- fullpvs %>% group_by(Species, Biome, Growth.Form, Ever.Decid) %>% summarise(TLP = -1*mean(TLP, na.rm=T),
                                                                                                   totRWCtlp = mean(totRWCtlp, na.rm=T),
                                                                                                   Po = mean(Po, na.rm=T),
                                                                                                   LMA = mean(LMA, na.rm=T),
                                                                                                   SWC = mean(SWC, na.rm=T),
                                                                                                   LMA_imputed = max(LMA_imputed),
                                                                                                   SWC_imputed = max(SWC_imputed),
                                                                                                   ApoFrac = mean(ApoFrac, na.rm=T),
                                                                                                   SWA = mean(SWA, na.rm=T),
                                                                                                   LWAtlp = mean(LWAtlp, na.rm=T),
                                                                                                   water_loss.m2 = mean(water_loss.m2, na.rm=T),
                                                                                                   CFT.g.m2 = mean(CFT.g.m2, na.rm=T),
  water_loss.g = mean(water_loss.g, na.rm=T),
                                                                                                   n_vals = n(), nDatasets=n_distinct(Dataset), Dataset = unique(Dataset)[1], T.S.L.H = unique(na.omit(T.S.L.H))[1]
                                                                                     )

fullspp <- fullspp[which(!is.na(fullspp$Species)),] 

# Some species have multiple values for E/D or Biome
  #  Celtis australis   Cordia collococca   Eucalyptus rubida        Oryza sativa 
  #                 2                   2                   2                   2 
  # Paullinia pinnata  Phaseolus vulgaris       Quercus rubra Tetracera volubilis 
  #                 2                   2                   2                   2 
  #   Trigonia rugosa   Triticum aestivum      Vitis vinifera 
  #                 2                   2                   3 
```




               
``` {r editing master}
# cleaning and adding in family info
# bin some of the low n biomes
fullspp$Biome[which(fullspp$Biome %in% c("Coastal","SemiDesert","Tundra"))] <- "Other"

#***NOTE: I switch the sign on TLP for ease of plotting!

# require(taxize)
# family.itis <- data.frame("db"=rep(NA, nrow(fullspp)),"query"=rep(NA, nrow(fullspp)), "family"=rep(NA, nrow(fullspp)))
# for( i in 1:nrow(fullspp)){
#   family.itis[i,] <- unlist(tax_name(fullspp$Species[i],get="family",db="itis"))
# } 
# # bonks on Larix decidua [183]
# family.itis$query[183] <- "Larix decidua"
# family.itis$family[183]="Pinaceae"
# 
# # second time bonks at 162
# # family.itis$query[162] <- "Polystichum munitum"
# # family.itis$family[162] <- "Dryopteridaceae"
# 
# #  for( i in 184:nrow(fullspp)){
# #    family.itis[i,] <- unlist(tax_name(fullspp$Species[i],get="family",db="itis"))
# #  }
# # 
# # family.itis1 <- family.itis
# # family.itis$query[183] <- "Polystichum munitum"
# # 
# # family.itis$query[183] <- "Polystichum munitum"
# # family.itis$family[162] <- "Dryopteridaceae"
# # for( i in 163:nrow(fullspp)){
# #   family.itis[i,] <- unlist(tax_name(fullspp$Species[i],get="family",db="itis"))
# # }
# 
# # fullspp$Species[162]
# # family.itis1$query <- fullspp$Species
# #write.csv(family.itis, "DerivedData/Family_lookuptable_ITIS_20240812.csv")
# missing <- family.itis1$query[which(is.na(family.itis1$family))]
# family.ncbi <- data.frame("db"=rep(NA, length(missing)),"query"=rep(NA, length(missing)), "family"=rep(NA, length(missing)))
# for(i in 1:length(missing)){
#   family.ncbi[i,] <- unlist(tax_name(missing[i],get="family",db="ncbi"))
# }
# # for(i in 19:length(missing)){
# #   family.ncbi[i,] <- unlist(tax_name(missing[i],get="family",db="ncbi"))
# # }
# # for(i in 27:length(missing)){
# #   family.ncbi[i,] <- unlist(tax_name(missing[i],get="family",db="ncbi"))
# #}
# 
# family.itis$family[which(is.na(family.itis$family))] <- family.ncbi$family[match(family.itis$query[which(is.na(family.itis$family))], family.ncbi$query)]
# 
# #write.csv(family.itis, "DerivedData/Family_lookuptable_ITIS-NCBI_20230802.csv")
# write.csv(family.itis, "DerivedData/Family_lookuptable_ITIS-NCBI_20240812.csv")

family.itis <- read.csv("DerivedData/Family_lookuptable_ITIS-NCBI_20240812.csv")

# add in Families to spp average and full dataset
fullspp$Family <- family.itis$family[match(fullspp$Species,family.itis$query)]
fullspp$Genus <- str_split_i(fullspp$Species," ", 1)
fullspp$Species_name <- str_split_i(fullspp$Species," ", -1)
fullpvs$Family <- family.itis$family[match(fullpvs$Species,family.itis$query)]
fullpvs$Genus <- str_split_i(fullpvs$Species," ", 1)
fullpvs$Species_name <- str_split_i(fullpvs$Species," ", -1)


# gymnosperm families
gymnos <- c("Araucariaceae",
"Cupressaceae",
"Gnetaceae",
"Pinaceae",
"Podocarpaceae",
"Taxaceae",
"Taxodiaceae",
"Zamiaceae")

fullspp$Gymno.Angio <- "Angiosperm"
fullspp$Gymno.Angio[which(fullspp$Family %in% gymnos)] <- "Gymnosperm"

fullpvs$Gymno.Angio <- "Angiosperm"
fullpvs$Gymno.Angio[which(fullpvs$Family %in% gymnos)] <- "Gymnosperm"

# Remove CFT and water_loss of Picea mariana because it's MAJ outlier
fullspp$CFT.g.m2[which(fullspp$Species=="Picea mariana")] <- NA
fullspp$water_loss.m2[which(fullspp$Species=="Picea mariana")] <- NA



fullfam <- fullspp %>% group_by(Family, Gymno.Angio) %>% summarise(TLP = -1*mean(TLP, na.rm=T),
                                                                                                   totRWCtlp = mean(totRWCtlp, na.rm=T),
                                                                                                   Po = mean(Po, na.rm=T),
                                                                                                   LMA = mean(LMA, na.rm=T),
                                                                                                   SWC = mean(SWC, na.rm=T),
                                                                                                   LMA_imputed = max(LMA_imputed),
                                                                                                   SWC_imputed = max(SWC_imputed),
                                                                                                   ApoFrac = mean(ApoFrac, na.rm=T),
                                                                                                   SWA = mean(SWA, na.rm=T),
                                                                                                   LWAtlp = mean(LWAtlp, na.rm=T),
                                                                                                   water_loss.m2 = mean(water_loss.m2, na.rm=T),
                                                                                                   CFT.g.m2 = mean(CFT.g.m2, na.rm=T),
  water_loss.g = mean(water_loss.g, na.rm=T),                                                       n_vals = n()  )





```

Still missing 50 families somehow, but will go back and fix those



``` {r missing metadata}

apply(fullspp, MARGIN=2, FUN = function(x){length(which(is.na(x)))})
```

# need:
38 Ever.Decid\
53 T.S.L.H\
50 family\
\
\


``` {r Clean edit data}

```


``` {r export derived datasets}

# full observations
write.csv(fullpvs, "DerivedData/FullObservations_Derived_20241001.csv")
# species averages
write.csv(fullspp, "DerivedData/SpeciesAverages_Derived_20241001.csv")



```
