---
title: "Exploration_and_Analysis"
author: "LDL Anderegg"
date: "10/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(dplyr)
library(reshape2)
mypal <- brewer.pal(n=8, "Set1")
palette(mypal)
```

## Exploring the implications of leaf PV curves for remote sensing of plant water content
Preliminary data exploration and analysis design



*** Analysis of summary parameters ****
Many of our datasets are just the summary parameters extracted from the leaf Pressure-Volume curves, rather than all of the raw measurements. From these, we're most interested in TLP (water potential at which leaf loses turgor and most leaf functioning stops) and RWC_tlp (the relative water content of the leaf at TLP)

```{r import data}
mb12raw <- read.csv("data/params/Bartlett2012ELE.csv", header=T) # data from Bartlett 2012 ELE
  # make clean version with all rows with both TLP and RWC_tlp values
mb12 <- mb12[which(mb12$TLP<0 & mb12$RWC_tlp>0),]
  # relevel things so that empty levels don't use all my colors
mb12$Biome <- relevel(mb12$Biome, ref="Temperate")
mb12$Biome <- relevel(mb12$Biome, ref="TropicalDry")

mb14wraw <- read.csv("data/params/Bartlett2014ELE_wild.csv") # data from Bartlett 2014 ELE wild plants
mb14wraw <- mb14w %>% rename(TLPwet=TLP_WET_MPa, TLPdry=TLP_DRY_MPa, RWC_tlpwet=RWC_total_wet_percent, RWC_tlpdry = RWC_total_dry_percent)
mb14w <- mb14wraw[which(mb14wraw$RWC_tlpwet>0),]
mb14c <- read.csv("data/params/Bartlett2014ELE_S3crop.csv")

qgar <- read.csv("data/params/Hahm_2018_Ecosphere_Quercusgarryana_20210115.csv")
  # a number of empty rows came in at the end, so kill those
qgar <- qgar[which(!is.na(qgar$Tree.ID)),]
  # relevel the 'rehydration' column so it doesn't have an empty NA level
qgar$Rehydration <- factor(qgar$Rehydration)
```

## We care about the actual raw Psi~100-RWC data, meaning we have to recalculate this from the PVcurve parameters
Psi_leaf is Psi_turgor + Psi_osmotic.

Psi_turgor = -1 * pi_o - mod_elast * R
and 
-1/Psi_osmotic = 1/pi_o - (1/pi_o)/(100-af) * R

pi_o = osmotic potential @ full turgor
af = apoplastic water fraction
R = 100-RWC

so
Psi_osmotic = -1/((1/pi_o) - R*(1/pi_o)/(100-af))



```{r translation functions, echo=FALSE}
Psi_turgor <- function(RWC, pi_o, mod_elast, af){
  RWCs <- (RWC-af*100)/(100-af*100) *100
  Rs <- (100-RWCs)
  Psi_turgor <- -1 * pi_o - mod_elast * Rs
  return(Psi_turgor)
} 



```



Exploratory visuals of four types of questions:
1) do biomes/land cover types differ?
2) how much does within-species seasonal adjustment matter?
3) how much does method (measuring rehydrated leaves vs measuring unrehydrated leaves) matter?
4) How appropriate is a linear fit for actual data?

``` {r quick plots RWC~TLP}

mypal1 <- brewer.pal(n=8, "Set1")
mypal1light <- paste0(mypal1, "66")
palette(mypal1light)


plot(TLP~RWC_tlp, mb12, col=Biome, xlim=c(68,100), ylim=c(min(mb12$TLP,na.rm=T),0), pch=16, cex=.8, xlab="%RWC")
mtext(side=3, "Global Meta-analysis")
for(i in 1:nrow(mb12)){
  lines(x=c(100,mb12$RWC_tlp[i]), y=c(0,mb12$TLP[i]), col=mb12$Biome[i])
}
legend('topleft', cex=.8, bty='n', legend = c("TropicalDry","Temperate"), pch=16, col=mypal1light[1:2])

plot(TLPdry~RWC_tlpdry, mb14w, col=mypal[1], xlim=c(60,100), ylim=c(min(mb14w$TLPdry,na.rm=T),0), pch=16, cex=.8, ylab="TLP", xlab="%RWC")
for(i in 1:nrow(mb14w)){
  lines(x=c(100,mb14w$RWC_tlpdry[i]), y=c(0,mb14w$TLPdry[i]), col=paste0(mypal[1],"33"))
}

points(TLPwet~RWC_tlpwet, mb14w, col=mypal[2], pch=16, cex=.8)
for(i in 1:nrow(mb14w)){
  lines(x=c(100,mb14w$RWC_tlpwet[i]), y=c(0,mb14w$TLPwet[i]), col=paste0(mypal[2],"33"))
}
arrows(x0=mb14w$RWC_tlpwet, x1=mb14w$RWC_tlpdry, y0=mb14w$TLPwet, y1=mb14w$TLPdry, length = .05)
legend('topleft', cex=.8, bty='n', legend = c("Dry Season","Wet Season"), pch=16, col=mypal1[1:2])
mtext("Seasonal Adjustment")


### Use the Quercus garyana dataset to look at how much rehydrating leaves before making PV curve measurements matters
plot(TLP~RWC_tlp, qgar, col=Rehydration, ylim=c(-5,0),xlim=c(.50,1.00), pch=16, ylab="TLP", xlab="%RWC")
for(i in 1:nrow(qgar)){
  lines(x=c(1,qgar$RWC_tlp[i]), y=c(0,qgar$TLP[i]), col=qgar$Rehydration[i])
}
legend('topleft', legend=c("unrehydrated", 'rehydrated'), col=mypal[c(1,2)], pch=16)
mtext("Method:Rehyd vs no rehyd")


### Appropriateness of linear fit when you have actual data
plot(WP~RWC, hf, ylab="Water Potential (MPa)", xlab="Relative Water Content", col=type)
legend('topleft', legend = c("Leaf","Leaf+Branch"), col=c(1,2), pch=1)
abline(h=-1.68, lty=2)
text(y=-2.1, x=3, "TLP\n(-1.68 MPa)")
for(i in unique(hf$sample)){
  lines(WP~RWC, hf[which(hf$sample==i),], col=type)
}
abline(lm(WP~RWC, hf[which(hf$WP> -1.68),]), col="black")
mtext("Linear Fit?")

```



**Analyses for Raw Data **
For data sources with raw RWC~Psi data, we don't have to back out the curve from the parameter values, and can instead calculate it directly from observations

``` {r load raw data sources}
## Holtzman et al. data from Harvard Forest
hf <- read.csv("data/raw_RWC-Psi/Holtzman_HarvardForest_rawPVcurves_20200304.csv"
)

plot(WP~R, hf, ylab="Water Potential (MPa)", xlab="100 - % Relative Water Content", col=type)
legend('bottomright', legend = c("Leaf","Leaf+Branch"), col=c(1,2), pch=1)
abline(h=-1.68, lty=2)
text(y=-2.1, x=3, "TLP\n(-1.68 MPa)")
abline(lm(WP~0 + R, hf[which(hf$WP> -1.68),]), col="grey") # add the regression line, forcing it through WP=0 @ R=0

```