---
title: "Exploration_and_Analysis"
author: "LDL Anderegg"
date: "10/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring the implications of leaf PV curves for remote sensing of plant water content
Preliminary data exploration and analysis design

```{r import data}
mb12 <- read.csv("data/params/Bartlett2012ELE.csv", header=T) # data from Bartlett 2012 ELE
mb14w <- read.csv("data/params/Bartlett2014ELE_wild.csv") # data from Bartlett 2014 ELE wild plants
mb14c <- read.csv("data/params/Bartlett2014ELE_S3crop.csv")

qgar <- read.csv("data/params/Hahm_2018_Ecosphere_Quercusgarryana_20210115.csv")
```

## We care about the actual raw Psi~100-RWC data, meaning we have to recalculate this from the PVcurve parameters
Psi_leaf is Psi_turgor + Psi_osmotic.

Psi_turgor = -1 * pi_o - mod_elast * R
and 
-1/Psi_osmotic = 1/pi_o - (1/pi_o)/(100-af) * R

pi_o = osmotic potential @ full turgor
af = apoplastic water fraction
R = 100-RWC

so
Psi_osmotic = -1/((1/pi_o) - R*(1/pi_o)/(100-af))



```{r translation functions, echo=FALSE}
Psi_turgor <- function(RWC, pi_o, mod_elast, af){
  RWCs <- (RWC-af*100)/(100-af*100) *100
  Rs <- (100-RWCs)
  Psi_turgor <- -1 * pi_o - mod_elast * Rs
  return(Psi_turgor)
} 



```



``` {r quick look at how much RWC_tlp varies}

hist(mb12$RWC_tlp, xlab="RWC at turgor loss point")
```



**Analyses for Raw Data **
For data sources with raw RWC~Psi data, we don't have to back out the curve from the parameter values, and can instead calculate it directly from observations

``` {r load raw data sources}
## Holtzman et al. data from Harvard Forest
hf <- read.csv("data/raw_RWC-Psi/Holtzman_HarvardForest_rawPVcurves_20200304.csv"
)

plot(WP~R, hf, ylab="Water Potential (MPa)", xlab="100 - % Relative Water Content", col=type)
legend('bottomright', legend = c("Leaf","Leaf+Branch"), col=c(1,2), pch=1)
abline(h=-1.68, lty=2)
text(y=-2.1, x=3, "TLP\n(-1.68 MPa)")
abline(lm(WP~0 + R, hf[which(hf$WP> -1.68),]), col="grey") # add the regression line, forcing it through WP=0 @ R=0

```