---
title: "Exploration_and_Analysis"
author: "LDL Anderegg"
date: "10/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(dplyr)
library(reshape2)
mypal <- brewer.pal(n=8, "Set1")
palette(mypal)
require(ggplot2)
```

## Exploring the implications of leaf PV curves for remote sensing of plant water content
Preliminary data exploration and analysis design



*** Analysis of summary parameters ****
Many of our datasets are just the summary parameters extracted from the leaf Pressure-Volume curves, rather than all of the raw measurements. From these, we're most interested in TLP (water potential at which leaf loses turgor and most leaf functioning stops) and RWC_tlp (the relative water content of the leaf at TLP)

```{r import data, include=F}
# data from Bartlett 2012 ELE
mb12raw <- read.csv("data/params/Bartlett2012ELE.csv", header=T) 
  # make clean version with all rows with both TLP and RWC_tlp values
mb12 <- mb12raw[which(mb12raw$TLP<0 & mb12raw$RWC_tlp>0),]
  # relevel things so that empty levels don't use all my colors
mb12$Biome <- factor(mb12$Biome)
mb12$Biome <- relevel(mb12$Biome, ref="Temperate")
mb12$Biome <- relevel(mb12$Biome, ref="TropicalDry")

# Bartlett 2014 ELE wild plants
mb14wraw <- read.csv("data/params/Bartlett2014ELE_wild.csv") # data from 
mb14wraw <- mb14wraw %>% rename(TLPwet=TLP_WET_MPa, TLPdry=TLP_DRY_MPa, RWC_tlpwet=RWC_total_wet_percent, RWC_tlpdry = RWC_total_dry_percent)
mb14w <- mb14wraw[which(mb14wraw$RWC_tlpwet>0),]
mb14c <- read.csv("data/params/Bartlett2014ELE_S3crop.csv")

# Data from Hahm 2018 Ecosphere
qgar <- read.csv("data/params/Hahm_2018_Ecosphere_Quercusgarryana_20210115.csv")
  # a number of empty rows came in at the end, so kill those
qgar <- qgar[which(!is.na(qgar$Tree.ID)),]
  # relevel the 'rehydration' column so it doesn't have an empty NA level
qgar$Rehydration <- factor(qgar$Rehydration)
qgar$LMA <- with(qgar, Dry_mass/Fresh_area)

## Data from Holtzman et al. 2021 from Harvard Forest
hf <- read.csv("data/raw_RWC-Psi/Holtzman_HarvardForest_rawPVcurves_20200304.csv"
)
hf$type <- factor(hf$type)
```

``` {r import contributed data, echo=F}
pvs <- read.csv("ContributedData/PV_for_RS_metaanalysis_20220716.csv", header=T)
pvs <- pvs[,-c(37:49)] # kills some extra columns that come in
# rename columns to be more code friendly
colnames(pvs) <- c("Contributor","Species","Treatment","Co.auth"
                   ,"Biome","Growth.Form","T.S.L","Ever.Decid"
                   ,"Wild.Garden","Life.Stage"
                   ,"TLP","TLS.sd","TLP.se"
                   , "totRWCtlp", "totRWCtlp.sd","totRWCtlp.se"
                   , "LMA", "SWC","CFT.mol.m2","Po","E","ApoFrac"
                   ,"sympRWCtlp","Psi.PD","Psi.min","Lat","Lon", "Elev"
                   ,"Site.Name","Rehydrated","Month","Year","OtherEnv"
                   ,"Methods.Notes","Ref","Other.Notes")
# fix LMA values that are in g cm2
pvs$LMA[which(pvs$LMA<1)] <- pvs$LMA[which(pvs$LMA<1)] * 100 * 100
# fix expanded Biomes
pvs$Biome[which(pvs$Biome=="TemperateConifer (but could be montane conifer…)")] <- "TemperateConifer"
# calculate the g of h20 per m2 of leaf area
pvs$sat.gwat.m2 <- pvs$SWC * pvs$LMA
# calculate the g of h20 per m2 leaf area at tlp
pvs$tlp.gwat.m2 <-pvs$sat.gwat.m2 * pvs$totRWCtlp /100
# calculate the water lost per m2 leaf area between saturation and tlp
pvs$water_loss.m2 <- with(pvs, sat.gwat.m2 - tlp.gwat.m2)
# turn capacitance into g instead of mol
pvs$CFT.g.m2 <- pvs$CFT.mol.m2 * 18.01528

# fix TLPs that are positive rather than negative
pvs$TLP[which(pvs$TLP>0)] <- -1 * pvs$TLP[which(pvs$TLP>0)]

```




``` {r Saturated water per mass vs area, include=F}
par(mar=c(8,5,2,1))
boxplot(SWC~Biome, pvs, las=2
        , xlab="", ylab=expression(paste("g water per g dry mass")))
mtext(text=paste("N=",length(which(pvs$SWC>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$SWC>0)]))), side=3,line = 0, adj = 0)

boxplot(sat.gwat.m2~Biome, pvs, las=2
        , xlab="", ylab=expression(paste("g water per ",m^2," leaf area")))
mtext(text=paste("N=",length(which(pvs$sat.gwat.m2>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$sat.gwat.m2>0)]))), side=3,line = 0, adj = 0)

ggplot(pvs, aes(x=SWC,y=sat.gwat.m2, col=Biome)) + geom_point() + xlab("Saturated Water Content") + ylab("g water per m2 leaf")


```
``` {r how much water can they lose, echo=F}
par(mar=c(8,5,2,1))
boxplot(water_loss.m2~Biome, pvs, las=2, ylim=c(0,200), ylab="g water lost\nfull sat to tlp, per m2 leaf", xlab="")
  #note: there's one value of ~480, which should be investigated
mtext(text=paste("N=",length(which(pvs$water_loss.m2>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)

boxplot(CFT.g.m2~Biome, pvs, las=2, ylab=expression(paste("g water lost ", MPa^-1, m^-2)), xlab="")
mtext(text=paste("N=",length(which(pvs$CFT.g.m2>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$CFT.g.m2>0)]))), side=3,line = 0, adj = 0)

```
TAKEHOME: conifers have a lot more water to lose per m2 leaf because they've got very high LMA and maybe fairly negative TLPs (difference between water loss per MPa and total water loss from saturation to TLP is the TLP itself).



``` {r plotting raw Psi per g curves, echo=F}
plot(TLP~sat.gwat.m2, pvs, type="n"
     , xlim=c(min(pvs$tlp.gwat.m2,na.rm=T), max(pvs$sat.gwat.m2,na.rm=T))
     , ylim=c(min(pvs$TLP, na.rm=T), 0)
     , ylab="Water Potential (MPa)"
     , xlab="g water per m2 leaf")
for(i in which(pvs$Species=="Sequoiadendron giganteum")){
  lines(x=c(pvs$sat.gwat.m2[i],pvs$tlp.gwat.m2[i]), y=c(0,pvs$TLP[i]), col="green")
}
for(i in 1:nrow(pvs)){
  lines(x=c(pvs$sat.gwat.m2[i],pvs$tlp.gwat.m2[i]), y=c(0,pvs$TLP[i]), col="#11111144")
}
mtext(text=paste("N=",length(which(pvs$water_loss.m2>0)), ";  # Species=",length(unique(pvs$Species[which(pvs$water_loss.m2>0)]))), side=3,line = 0, adj = 0)


```




OLD: Exploratory visuals of four types of questions:
1) do biomes/land cover types differ?
2) how much does within-species seasonal adjustment matter?
3) how much does method (measuring rehydrated leaves vs measuring unrehydrated leaves) matter?
4) How appropriate is a linear fit for actual data?

``` {r quick plots RWC~TLP}

mypal1 <- brewer.pal(n=8, "Set1")
mypal1light <- paste0(mypal1, "66")
palette(mypal1light)


plot(TLP~RWC_tlp, mb12, col=Biome, xlim=c(68,100), ylim=c(min(mb12$TLP,na.rm=T),0), pch=16, cex=.8, xlab="%RWC")
mtext(side=3, "Global Meta-analysis")
for(i in 1:nrow(mb12)){
  lines(x=c(100,mb12$RWC_tlp[i]), y=c(0,mb12$TLP[i]), col=mb12$Biome[i])
}
legend('topleft', cex=.8, bty='n', legend = c("TropicalDry","Temperate"), pch=16, col=mypal1light[1:2])

plot(TLPdry~RWC_tlpdry, mb14w, col=mypal[1], xlim=c(60,100), ylim=c(min(mb14w$TLPdry,na.rm=T),0), pch=16, cex=.8, ylab="TLP", xlab="%RWC")
for(i in 1:nrow(mb14w)){
  lines(x=c(100,mb14w$RWC_tlpdry[i]), y=c(0,mb14w$TLPdry[i]), col=paste0(mypal[1],"33"))
}

points(TLPwet~RWC_tlpwet, mb14w, col=mypal[2], pch=16, cex=.8)
for(i in 1:nrow(mb14w)){
  lines(x=c(100,mb14w$RWC_tlpwet[i]), y=c(0,mb14w$TLPwet[i]), col=paste0(mypal[2],"33"))
}
arrows(x0=mb14w$RWC_tlpwet, x1=mb14w$RWC_tlpdry, y0=mb14w$TLPwet, y1=mb14w$TLPdry, length = .05)
legend('topleft', cex=.8, bty='n', legend = c("Dry Season","Wet Season"), pch=16, col=mypal1[1:2])
mtext("Seasonal Adjustment")


### Use the Quercus garyana dataset to look at how much rehydrating leaves before making PV curve measurements matters
plot(TLP~RWC_tlp, qgar, col=Rehydration, ylim=c(-5,0),xlim=c(.50,1.00), pch=16, ylab="TLP", xlab="%RWC")
for(i in 1:nrow(qgar)){
  lines(x=c(1,qgar$RWC_tlp[i]), y=c(0,qgar$TLP[i]), col=qgar$Rehydration[i])
}
legend('topleft', legend=c("unrehydrated", 'rehydrated'), col=mypal[c(1,2)], pch=16)
mtext("Method:Rehyd vs no rehyd")


### Appropriateness of linear fit when you have actual data
plot(WP~RWC, hf, ylab="Water Potential (MPa)", xlab="Relative Water Content", col=type)
legend('topleft', legend = c("Leaf","Leaf+Branch"), col=c(1,2), pch=1)
abline(h=-1.68, lty=2)
text(y=-2.1, x=3, "TLP\n(-1.68 MPa)")
for(i in unique(hf$sample)){
  lines(WP~RWC, hf[which(hf$sample==i),], col=type)
}
abline(lm(WP~RWC, hf[which(hf$WP> -1.68),]), col="black")
mtext("Linear Fit?")

```















############## OLD OR ORPHAN CODE ##################



## We care about the actual raw Psi~100-RWC data, meaning we have to recalculate this from the PVcurve parameters
Psi_leaf is Psi_turgor + Psi_osmotic.

Psi_turgor = -1 * pi_o - mod_elast * R
and 
-1/Psi_osmotic = 1/pi_o - (1/pi_o)/(100-af) * R

pi_o = osmotic potential @ full turgor
af = apoplastic water fraction
R = 100-RWC

so
Psi_osmotic = -1/((1/pi_o) - R*(1/pi_o)/(100-af))



```{r translation functions, echo=FALSE}
Psi_turgor <- function(RWC, pi_o, mod_elast, af){
  RWCs <- (RWC-af*100)/(100-af*100) *100
  Rs <- (100-RWCs)
  Psi_turgor <- -1 * pi_o - mod_elast * Rs
  return(Psi_turgor)
} 



```







**Analyses for Raw Data **
For data sources with raw RWC~Psi data, we don't have to back out the curve from the parameter values, and can instead calculate it directly from observations

``` {r load raw data sources}

plot(WP~R, hf, ylab="Water Potential (MPa)", xlab="100 - % Relative Water Content", col=factor(type))
legend('bottomright', legend = c("Leaf","Leaf+Branch"), col=c(1,2), pch=1)
abline(h=-1.68, lty=2)
text(y=-2.1, x=3, "TLP\n(-1.68 MPa)")
abline(lm(WP~0 + R, hf[which(hf$WP> -1.68),]), col="grey") # add the regression line, forcing it through WP=0 @ R=0

```